---
title: "raw_count_data_generation"
output: html_document
date: "2022-10-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'missing-protein-project')
```




```{r}
library(tidyverse)
```




# import raw data file
```{r}
full_rawdata <- read_csv('data/nsaf_accession_new.csv', col_types = cols(
  gene = col_character(),
  accession = col_character(),
  length = col_double(),
  ipas = col_character(),
  description = col_character(),
  disease = col_character(),
  subtype = col_character(),
  note = col_character(),
  tce_mscnt = col_double(),
  # tce_mslen = col_double(),
  tce_nsaf = col_double(),
  media_mscnt = col_double(),
  # media_mslen = col_double(),
  media_nsaf = col_double(),
  media_ratio = col_double(),
  surface_mscnt = col_double(),
  # surface_mslen = col_double(),
  surface_nsaf = col_double(),
  surface_ratio = col_double(),
  nuclear_mscnt = col_double(),
  # nuclear_mslen = col_double(),
  nuclear_nsaf = col_double(),
  nuclear_ratio = col_double(),
  tee_mscnt = col_double(),
  # tee_mslen = col_double(),
  tee_nsaf = col_double(),
  # exosurface_mscnt = col_double(),
  # exosurface_mslen = col_double(),
  # exosurface_nsaf = col_double(),
  # exosurface_ratio = col_double()
  )
)
```

# import experiment meta data for samples
```{r}
experiment_metadata <- readxl::read_xlsx('data/sample_experiment_metadata.xlsx',
                                         sheet = 'Sheet1',
                                         skip = 2,
                                         col_names = c('ipas', 'description', 'type', 'organism', 'disease', 'subtype', 'note',
                                                       'S',	'N',	'S',	'N',	 'S',	'N',	'S',	'N',	'S',	'N',	'S',	'N',	'S',	'N',
                                                       'issue', 'mysql', 'mongo'))

# cell_line <- experiment_metadata %>% filter(type == 'Cell Line', mysql == '1701') %>% select(ipas) %>%  pull %>% sort()
```



```{r}
sort(unique(full_rawdata$ipas))
```


# import normalized data
```{r}
SpC_TCE <- read.delim("data/CellLineCompartment_TCE_SumTickNormalization.tsv", sep = '\t', skip=2, row.names = 'Accession')
```


```{r}
SpC_surface <- read.delim("data/CellLineCompartment_Surface_SumTickNormalization.tsv",  sep = '\t', skip=2, row.names = 'Accession')
SpC_exosome <- read.csv("data/CellLineCompartment_Exosome_SumTickNormalization.csv", skip=2, row.names = 'Accession')
SpC_media <- read.csv("data/CellLineCompartment_Media_SumTickNormalization.csv", skip=2, row.names = 'Accession')
SpC_nuclear <- read.delim("data/CellLineCompartment_Nuclear_SumTickNormalization.csv", sep = '\t', skip=2, row.names = 'Accession')
SpCs <- list(TCE=SpC_TCE, surface=SpC_surface, exosome=SpC_exosome, media=SpC_media, nuclear=SpC_nuclear)
```


# extract IPAS numbers in each component dataset
```{r}
extract_IPAS <- function(matrix) {
  str_extract(colnames(matrix), '(?<=\\.)IP.+?(?=\\.)')
}

IPAS_list <- lapply(SpCs, extract_IPAS)
# for(i in seq(1, 300, 50)) {
#   print(paste0(IPAS_list[i:(i+50)], collapse = ', '))
# }

```


```{r}
generating_raw_count <- function(IPAS, count_column, keep_isoform=FALSE) {
  raw_count <- full_rawdata %>% 
   dplyr::select(accession, ipas, !!count_column) %>%
   filter(ipas %in% IPAS) %>%
   pivot_wider(names_from = ipas, values_from = !!count_column) %>% 
   replace(is.na(.), 0)

  raw_count_mat <- as.matrix(raw_count[-1])
  rownames(raw_count_mat) <- pull(raw_count[1])
  raw_count_mat <- raw_count_mat[rowSums(raw_count_mat) > 0, ]
  raw_count_mat <- raw_count_mat[, colSums(raw_count_mat) > 0 ]
  
  if(!keep_isoform) {
    raw_count_mat <- remove_isoforms(raw_count_mat)
  }
  
  cat(count_column, ': ',
      'number of col', ncol(raw_count_mat), '\n',
      'number of row', nrow(raw_count_mat), '\n')
  
  return(raw_count_mat)
}

# example
# generating_raw_count(IPAS_list$TCE, count_column = 'tce_mscnt')
```



```{r}
raw_count_mats <- mapply(generating_raw_count, IPAS_list[-which(names(IPAS_list) == 'exosome')], c('tce_mscnt', 'surface_mscnt', 'media_mscnt', 'nuclear_mscnt'))
```
# check overlap
```{r}
IPAS_list$TCE[which(!IPAS_list$TCE %in% unique(full_rawdata$ipas))]
IPAS_list$surface[which(!IPAS_list$surface %in% unique(full_rawdata$ipas))]
IPAS_list$media[which(!IPAS_list$media %in% unique(full_rawdata$ipas))]
IPAS_list$nuclear[which(!IPAS_list$nuclear %in% unique(full_rawdata$ipas))]
```


```{r}
# head(raw_count_mats$TCE)
```

# check distribution !take long time to run
```{r}
# boxplot(as.data.frame(raw_count_mats$TCE))
```


# sort column by column names
```{r}
sort_by_column <- function(mat) {
  return(mat[, sort(colnames(mat))])
}

raw_count_mats <- lapply(raw_count_mats, sort_by_column)

# raw_count_mats$TCE <- raw_count_mats$TCE[ ,sort(colnames(raw_count_mats$TCE))]
```


# Approach 1
# combine isoforms by summing them
```{r}
# rownames(raw_count_mats$TCE) <- str_extract(rownames(raw_count_mats$TCE), pattern = '[^-]+')


combine_isoforms <- function(mat) {
  tbl <- mat %>% as_tibble(rownames = 'accession') %>% 
    mutate(accession=str_extract(accession, pattern = '[^-]+')) %>%
    group_by(accession) %>% 
    summarise_all(sum)
  
  out <- as.matrix(tbl[-1])
  # need to use pull to make it as vector
  rownames(out) <- pull(tbl[1])
  return(out)
}

# combine_isoforms(raw_count_mats$TCE)

# raw_count_mats_iso_combined <- lapply(raw_count_mats, combine_isoforms)

```


# Approach 2
# only keep canonical form (recommended by Jennifer) to avoid double-count
```{r}

remove_isoforms <- function(mat) {
  tbl <- mat %>% as_tibble(rownames = 'accession') %>% 
    mutate('is_isoform'=str_detect(accession, pattern = '.+-.+')) %>% 
    filter(is_isoform == FALSE) %>% 
    select(-c('is_isoform'))
  
  out <- as.matrix(tbl[-1])
  # need to use pull to make it as vector
  rownames(out) <- pull(tbl[1])
  return(out)
}

raw_count_mats_iso_removed <- lapply(raw_count_mats, remove_isoforms)
```


# export raw count data
```{r}
# helper function write data to provided path
write_to_csv <- function(mat, path) {
  write.csv(as.data.frame(mat, row.names = rownames(mat)), path)
}

write_to_csv(raw_count_mats_iso_removed$TCE, 'data/cell line raw data/tce_raw_count.csv')
write_to_csv(raw_count_mats_iso_removed$surface, 'data/cell line raw data/surface_raw_count.csv')
write_to_csv(raw_count_mats_iso_removed$media, 'data/cell line raw data/media_raw_count.csv')
write_to_csv(raw_count_mats_iso_removed$nuclear, 'data/cell line raw data/nuclear_raw_count.csv')
```



# generate omics metadata
# do the conversion using biomaRt
the result is better than using org.Hs.eg.db
no missing gene symbol

! there are some Ensembl gene id cannot be found by biomart
https://support.bioconductor.org/p/111608/


<!-- 15388 unmatched ensembl id -->
<!-- ```{r} -->
<!-- annots <- AnnotationDbi::select(org.Hs.eg.db::org.Hs.eg.db, keys=rownames(count_TPM_mat),  -->
<!--                 columns="SYMBOL", keytype="ENSEMBL") -->
<!-- ``` -->



```{r eval=FALSE, include=FALSE}
load('cache/count_TPM.rda')

mart <- biomaRt::useDataset("hsapiens_gene_ensembl", biomaRt::useMart("ensembl"))
dict_ensembl_to_symbol <- biomaRt::getBM(filters= "ensembl_gene_id", attributes= c("ensembl_gene_id","hgnc_symbol", "start_position","end_position"), values=rownames(count_TPM_mat), mart= mart)
dict_ensembl_to_symbol$gene_length=dict_ensembl_to_symbol$end_position - dict_ensembl_to_symbol$start_position
dict_ensembl_to_symbol[dict_ensembl_to_symbol == ''] <- NA
save(dict_ensembl_to_symbol, file='cache/dict_ensemble_to_symbol.rda')
```


```{r}
load('cache/dict_ensemble_to_symbol.rda')
# dict_ensembl_to_symbol
```

```{r eval=FALSE, include=FALSE}
accession_metadata <- full_rawdata %>% dplyr::select(gene, accession, length) %>% unique()
load('cache/dict_ensemble_to_symbol.rda')
omics_metadata <- dict_ensembl_to_symbol %>% full_join(accession_metadata, by = c('hgnc_symbol'='gene')) %>% dplyr::select(-c(start_position, end_position)) %>% dplyr::select(accession, hgnc_symbol, ensembl_gene_id, gene_length, length)
colnames(omics_metadata) <- c('accession_id', 'gene_symbol', 'ensembl_gene_id', 'gene_length', 'protein_length')
head(omics_metadata)
write_csv(omics_metadata, file = 'data/support-data/omics_metadata.csv')


omics_metadata <- read_csv('data/support-data/omics_metadata.csv')
protein_length <- omics_metadata %>% dplyr::select(accession_id, protein_length) %>% unique() %>% na.omit()
protein_length_df <- data.frame(length=protein_length$protein_length, row.names = protein_length$accession_id)
save(protein_length_df, file = 'cache/protein_length.rda')
```
```{r}
load('cache/protein_length.rda')
```




```{r}
nasf_mat <- lapply(raw_count_mats_iso_removed, nsaf, protein_length=protein_length_df, per_count=10e6)
```


# export nasf data
```{r}
write_to_csv(nasf_mat$TCE, 'data/cell line raw data/nasf_tce.csv')
write_to_csv(nasf_mat$surface, 'data/cell line raw data/nasf_surface.csv')
write_to_csv(nasf_mat$media, 'data/cell line raw data/nasf_media.csv')
write_to_csv(nasf_mat$nuclear, 'data/cell line raw data/nasf_nuclear.csv')
```


# only keep when count >= 2
Should I filter count >=2 then calculate nsaf?
```{r}
filter_greater_than_two <- function(count_mat) {
  count_mat[count_mat <= 1] <- 0
  return(count_mat)
}

raw_count_mats_iso_removed_2 <- lapply(raw_count_mats_iso_removed, filter_greater_than_two)
mapply(write_to_csv, raw_count_mats_iso_removed_2, path=list('data/cell line raw data/tce_raw_count_2.csv', 
                                                          'data/cell line raw data/surface_raw_count_2.csv',
                                                          'data/cell line raw data/media_raw_count_2.csv',
                                                          'data/cell line raw data/nuclear_raw_count_2.csv'))

nasf_mat <- lapply(raw_count_mats_iso_removed_2, nsaf, protein_length=protein_length_df, per_count=10e6)

write_to_csv(nasf_mat$TCE, 'data/cell line raw data/nasf_tce_2.csv')
write_to_csv(nasf_mat$surface, 'data/cell line raw data/nasf_surface_2.csv')
write_to_csv(nasf_mat$media, 'data/cell line raw data/nasf_media_2.csv')
write_to_csv(nasf_mat$nuclear, 'data/cell line raw data/nasf_nuclear_2.csv')
```







# filter with missing proteins
```{r}
missing_protein_df <- readxl::read_xlsx('data/support data/PE2-5.xlsx')
# missing_protein_df <- missing_protein_df %>% filter(PE != 'Uncertain')
save(missing_protein_df, file='cache/missing_protein.rda')
```



```{r}
nasf_mat_missing_protein <- lapply(nasf_mat, SpC_List, annotation=NULL, NA_substitution = NULL, proteins_filter = missing_protein_df$Accession)
```

# export nasf data with only missing protein
```{r}
write_to_csv(nasf_mat_missing_protein$TCE$matrix, 'data/nasf/nasf_tce_missing_protein.csv')
write_to_csv(nasf_mat_missing_protein$surface$matrix, 'data/nasf/nasf_surface_missing_protein.csv')
write_to_csv(nasf_mat_missing_protein$media$matrix, 'data/nasf/nasf_media_missing_protein.csv')
write_to_csv(nasf_mat_missing_protein$nuclear$matrix, 'data/nasf/nasf_nuclear_missing_protein.csv')
```


# generate annotation data for all samples in the nasf_accession file
```{r}

# create annotation for IPAS
IPAS_annotation <- full_rawdata %>% select(ipas, disease, subtype) %>% unique() %>% arrange(disease, subtype)
write_csv(IPAS_annotation, file = 'data/support data/IPAS_annotation.csv')
```


# generate primary cell raw count and NSAF data
```{r}
# since IP0972_1701 doesn't have corresponding MS count data, we won't keep it in the MS count data
IPAS_in_RNA_Seq <- paste0(c('IP7103', 'IP0995', 'IP0982', 'IP0999', 'IP7100', 'IP0993', 'IP0981', 'IP7105'), '_1701')
gastric_primary_cell_count_mat <- generating_raw_count(IPAS = IPAS_in_RNA_Seq, count_column = 'tce_mscnt', keep_isoform=FALSE)
```




```{r}
gastric_primary_cell_nsaf_mat <- nsaf(gastric_primary_cell_count_mat, protein_length_df, per_count = 10e6, na_fill=0)
gastric_primary_cell_nsaf_mat <- gastric_primary_cell_nsaf_mat[rowSums(gastric_primary_cell_nsaf_mat > 0) > 0, ]
head(gastric_primary_cell_nsaf_mat)
```


## export data
```{r}
write_to_csv(gastric_primary_cell_count_mat, 'data/gastric-primary-cell/gastric_primary_cell_count.csv')
write_to_csv(gastric_primary_cell_nsaf_mat, 'data/gastric-primary-cell/gastric_primary_cell_nsaf.csv')
```



# only keep count >= 2
```{r}
gastric_primary_cell_count_mat[gastric_primary_cell_count_mat <= 1] <- 0
gastric_primary_cell_nsaf_mat <- nsaf(gastric_primary_cell_count_mat, protein_length_df, per_count = 10e6, na_fill=0)
# fill those less than 2 spectral count read as 0 and filter proteins that are 0 for all samples
gastric_primary_cell_nsaf_mat <- gastric_primary_cell_nsaf_mat[rowSums(gastric_primary_cell_nsaf_mat > 0) > 0, ]
gastric_primary_cell_count_mat <- gastric_primary_cell_count_mat[rowSums(gastric_primary_cell_count_mat > 0) > 0, ]
write_to_csv(gastric_primary_cell_count_mat, 'data/gastric-primary-cell/gastric_primary_cell_count_2.csv')
write_to_csv(gastric_primary_cell_nsaf_mat, 'data/gastric-primary-cell/gastric_primary_cell_nsaf_2.csv')
```




```{r}
metadata <- read_csv('data/sample_experiment_metadata.csv', skip = 1)
metadata <- metadata %>% select(c('IPASx', 'Descriptionx', 'TypeSelect valueCell LinePrimary CellStem CellTissuex', 'OrganismSelect valueHumanHumanMouseMousex', 'DiseaseSelect valueBreastColonGastricGliomaLeukemiaLungLymphoblastMelanomaNeuroblastomaOvarianPancreaticProstatex',
                      'SubTypeSelect valueALLAMLAdenoAscitiesBasalBrennerTumorCLLCMLCarcinomaControlExocrineAdenoGlioblastomaHER2LuminalA/BLungSmallCellMDSMesenchymalMetastaticNSCLCNormalPDACTNBCXenox',
                      'MySQLx'))
```


```{r}
colnames(metadata) <- c('IPAS', 'description', 'type', 'organism', 'disease', 'subtype', 'MySQL')
```

```{r}
metadata_primary_cell <- metadata %>% filter(type == 'Primary Cell' & organism == 'Human' & MySQL == '1701') %>% select(IPAS, disease, subtype) %>% filter(!str_detect(IPAS, '_'))
```



```{r}
IPAS_primary_cell <- paste0(str_replace(metadata_primary_cell$IPAS, 'IPAS', 'IP') , '_1701')
primary_cell_count_mat <- generating_raw_count(IPAS = IPAS_primary_cell, count_column = 'tce_mscnt', keep_isoform=FALSE)
```

# only keep cell value with spectral count >=2
```{r}
primary_cell_count_mat[primary_cell_count_mat  <= 1] <- 0
primary_cell_nsaf_mat <- nsaf(primary_cell_count_mat, protein_length_df, per_count = 10e6, na_fill=0)
# remove row that are all 0
primary_cell_nsaf_mat <- primary_cell_nsaf_mat[rowSums(primary_cell_nsaf_mat > 0) > 0, ]
# head(primary_cell_nsaf_mat)
```


```{r}
write.csv(as.data.frame(primary_cell_nsaf_mat, row.names = rownames(primary_cell_nsaf_mat)), 'data/nasf/nasf_primary_cell_2.csv')
```

