---
title: "Missing protein analysis"
output: html_document
---

# FDR distribution
```{r echo=TRUE, warning=FALSE}
## let read the required libraries
rm(list = ls())
library(readxl)
library(pROC)
library(tidyr)
library(dplyr)

## let's set the directory
setwd("~/Box Sync/Missing protein project/")

## read the data first 
df_missing <- read_xlsx("./Data/missingprotein_FPR.xlsx")
df_missing$protein.falsePositiveRate = df_missing$protein.falsePositiveRate
df_nonmissing <- read.csv("./Data/FPR_nonmissing.csv")
df_nonmissing_distr <- sample(df_nonmissing$protein.falsePositiveRate,60000)


## density plot

P1 <- density(df_missing$protein.falsePositiveRate)
P2 <- density(df_nonmissing_distr)


xlim <- range(P1$x,P2$x)
ylim <- range(0,P1$y,P2$y)

plot(P2, xlim = c(0,2), ylim = ylim, xlab = 'FPR %',
     main = 'Distribution of FPR', 
     panel.first = grid())
polygon(P1, density = -1, col = scales::alpha('orange',0.6))
polygon(P2, density = -1, col = scales::alpha('orchid',0.4))

# polygon(P3, density = -1, col = scales::alpha('turquoise1',0.5))

legend('top',c('Missing proteins','Regular proteins'),
       fill = c(scales::alpha('orange',0.5), scales::alpha('orchid',0.5),col = scales::alpha('turquoise1',0.5)), bty = 'n',
       border = NA, cex = 1.2)
```


# TCE Heatmap
```{r echo=TRUE, warning=FALSE}

## let read the required libraries
rm(list = ls())

library(tidyr)
library(dplyr)
library(pheatmap)
library(RColorBrewer)
# library(heatmap.plus)
library(viridis)
library(dendextend)

## let's set the directory
# setwd("~/Box Sync/Missing protein project/")

## read the data first
df_missing_normalized <- read.csv("data/CellLineCompartment_TCE_SumTickNormalization.tsv",sep = "\t", skip = 2,row.names = 1)
df_missing_normalized_all <- read.csv("data/CellLineCompartment_TCE_SumTickNormalization.tsv",sep = "\t", header = F)

## annotation column
col_annotation <- data.frame("Cancer type"= t(df_missing_normalized_all[1,2:dim(df_missing_normalized_all)[2]]))

col_annotation$X2 <- as.character(col_annotation$X1)
col_annotation$X2[grep("Breast",col_annotation$X1)] = ("Breast")
col_annotation$X2[grep("Leukemia",col_annotation$X1)] = ("Leukemia")
col_annotation <- data.frame(col_annotation[,-c(1)])
rownames(col_annotation) <- colnames(df_missing_normalized)
colnames(col_annotation) <- c("Simplified Cancer type")


### log2 scaling
df_missing_normalized_matrix <- as.matrix(df_missing_normalized)
df_missing_normalized_matrix[is.na(df_missing_normalized_matrix)] = 0
df_missing_normalized_matrix[df_missing_normalized_matrix!=0] = log10(df_missing_normalized_matrix[df_missing_normalized_matrix!=0])

# ### z score normaliztion instead
# df_missing_normalized_matrix <- as.matrix(df_missing_normalized)
# df_missing_normalized_matrix[is.na(df_missing_normalized_matrix)] = 0
# df_missing_normalized_matrix <- apply(df_missing_normalized_matrix, 2, scale)

### heatmap
# pheatmap(df_missing_normalized_matrix,
#          #fontsize_col = 5, 
#          #fontsize_row = 4,
#          show_rownames = F,
#          show_colnames = F,
#          main = "Total Cell Extract", 
#          cluster_cols = F,
#          cluster_rows = F,
#          annotation_col = col_annotation)

#### hmmm! well it is so spares, let's focus on those being repeated

# only leave those spread apart? 
df_missing_normalized_matrix_frequestnidentified <- df_missing_normalized_matrix[apply(df_missing_normalized_matrix, 1, function(c) sd(c)>0.1),]

# at least two condtions have non-missing values
df_missing_normalized_matrix_frequestnidentified <- df_missing_normalized_matrix_frequestnidentified[apply(df_missing_normalized_matrix_frequestnidentified, 1, function(c) sum(c!=0)>1), ]




# mat_colors <- list(group = brewer.pal(sum(!duplicated(col_annotation$`Simplifeid Cancer type`)), "Set3"))

nb.cols <- sum(!duplicated(col_annotation$`Simplifeid Cancer type`))
Var1 <- colorRampPalette(brewer.pal(8, "Pastel1"))(nb.cols)
names(Var1) <- unique(col_annotation$`Simplifeid Cancer type`)
anno_colors <- list(Var1 = Var1)


quantile_breaks <- function(xs, n = 10) {
  breaks <- quantile(xs, probs = seq(0, 1, length.out = n))
  breaks[!duplicated(breaks)]
}

mat_breaks <- quantile_breaks(df_missing_normalized_matrix_frequestnidentified, n = 200)

breaksList = seq(0, 2, by = 0.05)
pheatmap(df_missing_normalized_matrix_frequestnidentified,
         #fontsize_col = 5, 
         fontsize_row = 3,
         # cutree_rows = 6,
         color = colorRampPalette(colors = c("black", "yellow", "darkorange"))(length(breaksList)),
         # color = colorRampPalette(colors = c("black", "yellow", "darkorange"))(length(mat_breaks)),
         # breaks = mat_breaks,
         show_rownames = T,
         show_colnames = F,
         main = "Total Cell Extract", 
         cluster_cols = F,
         cluster_rows = T,
         annotation_colors = anno_colors,
         annotation_col = col_annotation
         )

write.csv(df_missing_normalized_matrix_frequestnidentified, "output/TCE_abundant_missingpteinst.csv")
```



# Surface Heatmap
```{r echo=TRUE, warning=FALSE}

## let read the required libraries
rm(list = ls())

library(tidyr)
library(dplyr)
library(pheatmap)
library(RColorBrewer)
# library(heatmap.plus)
library(viridis)
library(dendextend)

## let's set the directory
# setwd("~/Box Sync/Missing protein project/")

## read the data first
df_missing_normalized <- read.csv("./Data/CellLineCompartment_Surface_SumTickNormalization.csv",sep = "\t", skip = 2,row.names = 1)
df_missing_normalized_all <- read.csv("./Data/CellLineCompartment_Surface_SumTickNormalization.csv",sep = "\t", header = F)

## annotation column
col_annotation <- data.frame("Cancer type"= t(df_missing_normalized_all[1,2:dim(df_missing_normalized_all)[2]]))
rownames(col_annotation) <- colnames(df_missing_normalized)
colnames(col_annotation) <- c("Cancer type")

### log2 scaling
df_missing_normalized_matrix <- as.matrix(df_missing_normalized)
df_missing_normalized_matrix[is.na(df_missing_normalized_matrix)] = 0
df_missing_normalized_matrix[df_missing_normalized_matrix!=0] = log10(df_missing_normalized_matrix[df_missing_normalized_matrix!=0])

# ### z score normaliztion instead
# df_missing_normalized_matrix <- as.matrix(df_missing_normalized)
# df_missing_normalized_matrix[is.na(df_missing_normalized_matrix)] = 0
# df_missing_normalized_matrix <- apply(df_missing_normalized_matrix, 2, scale)

### heatmap
pheatmap(df_missing_normalized_matrix,
         #fontsize_col = 5, 
         #fontsize_row = 4,
         show_rownames = F,
         show_colnames = F,
         main = "Surface", 
         cluster_cols = F,
         cluster_rows = F,
         annotation_col = col_annotation)

#### hmmm! well it is so spares, let's focus on those being repeated
df_missing_normalized_matrix_frequestnidentified <- df_missing_normalized_matrix[apply(df_missing_normalized_matrix, 1, function(c) sum(c!=0)>4),]

mat_colors <- list(group = brewer.pal(9, "Set1"))

nb.cols <- 17
Var1 <- colorRampPalette(brewer.pal(8, "Reds"))(nb.cols)
names(Var1) <- unique(col_annotation$`Cancer type`)
anno_colors <- list(Var1 = Var1)

breaksList = seq(0, 2, by = 0.05)
pheatmap(df_missing_normalized_matrix_frequestnidentified,
         #fontsize_col = 5, 
         fontsize_row = 6,
         # cutree_cols = 17,
         color = colorRampPalette(colors = c("black", "yellow", "darkorange"))(length(breaksList)),
         show_rownames = T,
         show_colnames = F,
         main = "Surface", 
         cluster_cols = F,
         cluster_rows = T,
         annotation_col = col_annotation)
         # annotation_colors = anno_colors)


```



# Media Heatmap
```{r echo=TRUE, warning=FALSE}

## let read the required libraries
rm(list = ls())

library(tidyr)
library(dplyr)
library(pheatmap)
library(RColorBrewer)
# library(heatmap.plus)
library(viridis)
library(dendextend)

## let's set the directory
# setwd("~/Box Sync/Missing protein project/")

## read the data first
df_missing_normalized <- read.csv("./Data/CellLineCompartment_Media_SumTickNormalization.csv", skip = 2,row.names = 1)
df_missing_normalized_all <- read.csv("./Data/CellLineCompartment_Media_SumTickNormalization.csv", header = F)

## annotation column
col_annotation <- data.frame("Cancer type"= t(df_missing_normalized_all[1,2:dim(df_missing_normalized_all)[2]]))
rownames(col_annotation) <- colnames(df_missing_normalized)
colnames(col_annotation) <- c("Cancer type")

### log2 scaling
df_missing_normalized_matrix <- as.matrix(df_missing_normalized)
df_missing_normalized_matrix[is.na(df_missing_normalized_matrix)] = 0
df_missing_normalized_matrix[df_missing_normalized_matrix!=0] = log10(df_missing_normalized_matrix[df_missing_normalized_matrix!=0])

# ### z score normaliztion instead
# df_missing_normalized_matrix <- as.matrix(df_missing_normalized)
# df_missing_normalized_matrix[is.na(df_missing_normalized_matrix)] = 0
# df_missing_normalized_matrix <- apply(df_missing_normalized_matrix, 2, scale)

### heatmap
pheatmap(df_missing_normalized_matrix,
         #fontsize_col = 5, 
         #fontsize_row = 4,
         show_rownames = F,
         show_colnames = F,
         main = "Media", 
         cluster_cols = F,
         cluster_rows = F,
         annotation_col = col_annotation)

#### hmmm! well it is so spares, let's focus on those being repeated
df_missing_normalized_matrix_frequestnidentified <- df_missing_normalized_matrix[apply(df_missing_normalized_matrix, 1, function(c) sum(c!=0)>4),]

mat_colors <- list(group = brewer.pal(9, "Set1"))

nb.cols <- 17
Var1 <- colorRampPalette(brewer.pal(8, "Reds"))(nb.cols)
names(Var1) <- unique(col_annotation$`Cancer type`)
anno_colors <- list(Var1 = Var1)

breaksList = seq(0, 2, by = 0.05)
pheatmap(df_missing_normalized_matrix_frequestnidentified,
         #fontsize_col = 5, 
         fontsize_row = 6,
         # cutree_cols = 17,
         color = colorRampPalette(colors = c("black", "yellow", "darkorange"))(length(breaksList)),
         show_rownames = T,
         show_colnames = F,
         main = "Media", 
         cluster_cols = F,
         cluster_rows = T,
         annotation_col = col_annotation)
         # annotation_colors = anno_colors)


```



# Nuclear Heatmap
```{r echo=TRUE, warning=FALSE}

## let read the required libraries
rm(list = ls())

library(tidyr)
library(dplyr)
library(pheatmap)
library(RColorBrewer)
# library(heatmap.plus)
library(viridis)
library(dendextend)

## let's set the directory
# setwd("~/Box Sync/Missing protein project/")

## read the data first
df_missing_normalized <- read.csv("./Data/CellLineCompartment_Nuclear_SumTickNormalization.csv",sep = "\t", skip = 2,row.names = 1)
df_missing_normalized_all <- read.csv("./Data/CellLineCompartment_Nuclear_SumTickNormalization.csv",sep = "\t", header = F)

## annotation column
col_annotation <- data.frame("Cancer type"= t(df_missing_normalized_all[1,2:dim(df_missing_normalized_all)[2]]))
rownames(col_annotation) <- colnames(df_missing_normalized)
colnames(col_annotation) <- c("Cancer type")

### log2 scaling
df_missing_normalized_matrix <- as.matrix(df_missing_normalized)
df_missing_normalized_matrix[is.na(df_missing_normalized_matrix)] = 0
df_missing_normalized_matrix[df_missing_normalized_matrix!=0] = log10(df_missing_normalized_matrix[df_missing_normalized_matrix!=0])

# ### z score normaliztion instead
# df_missing_normalized_matrix <- as.matrix(df_missing_normalized)
# df_missing_normalized_matrix[is.na(df_missing_normalized_matrix)] = 0
# df_missing_normalized_matrix <- apply(df_missing_normalized_matrix, 2, scale)

### heatmap
pheatmap(df_missing_normalized_matrix,
         #fontsize_col = 5, 
         #fontsize_row = 4,
         show_rownames = F,
         show_colnames = F,
         main = "Nucleas", 
         cluster_cols = F,
         cluster_rows = F,
         annotation_col = col_annotation)

#### hmmm! well it is so spares, let's focus on those being repeated
df_missing_normalized_matrix_frequestnidentified <- df_missing_normalized_matrix[apply(df_missing_normalized_matrix, 1, function(c) sum(c!=0)>4),]

mat_colors <- list(group = brewer.pal(9, "Set1"))

nb.cols <- 17
Var1 <- colorRampPalette(brewer.pal(8, "Reds"))(nb.cols)
names(Var1) <- unique(col_annotation$`Cancer type`)
anno_colors <- list(Var1 = Var1)

breaksList = seq(0, 2, by = 0.05)
pheatmap(df_missing_normalized_matrix_frequestnidentified,
         #fontsize_col = 5, 
         fontsize_row = 6,
         # cutree_cols = 17,
         color = colorRampPalette(colors = c("black", "yellow", "darkorange"))(length(breaksList)),
         show_rownames = T,
         show_colnames = F,
         main = "Nucleas", 
         cluster_cols = F,
         cluster_rows = T,
         annotation_col = col_annotation)
         # annotation_colors = anno_colors)


```




# Exosome Heatmap
```{r echo=TRUE, warning=FALSE}

## let read the required libraries
rm(list = ls())

library(tidyr)
library(dplyr)
library(pheatmap)
library(RColorBrewer)
# library(heatmap.plus)
library(viridis)
library(dendextend)

## let's set the directory
# setwd("~/Box Sync/Missing protein project/")

## read the data first
df_missing_normalized <- read.csv("./Data/CellLineCompartment_Exosome_SumTickNormalization.csv", skip = 2,row.names = 1)
df_missing_normalized_all <- read.csv("./Data/CellLineCompartment_Exosome_SumTickNormalization.csv", header = F)

## annotation column
col_annotation <- data.frame("Cancer type"= t(df_missing_normalized_all[1,2:dim(df_missing_normalized_all)[2]]))
rownames(col_annotation) <- colnames(df_missing_normalized)
colnames(col_annotation) <- c("Cancer type")

### log2 scaling
df_missing_normalized_matrix <- as.matrix(df_missing_normalized)
df_missing_normalized_matrix[is.na(df_missing_normalized_matrix)] = 0
df_missing_normalized_matrix[df_missing_normalized_matrix!=0] = log10(df_missing_normalized_matrix[df_missing_normalized_matrix!=0])

# ### z score normaliztion instead
# df_missing_normalized_matrix <- as.matrix(df_missing_normalized)
# df_missing_normalized_matrix[is.na(df_missing_normalized_matrix)] = 0
# df_missing_normalized_matrix <- apply(df_missing_normalized_matrix, 2, scale)

### heatmap
pheatmap(df_missing_normalized_matrix,
         #fontsize_col = 5, 
         #fontsize_row = 4,
         show_rownames = F,
         show_colnames = F,
         main = "Exosome", 
         cluster_cols = F,
         cluster_rows = F,
         annotation_col = col_annotation)

#### hmmm! well it is so spares, let's focus on those being repeated
df_missing_normalized_matrix_frequestnidentified <- df_missing_normalized_matrix[apply(df_missing_normalized_matrix, 1, function(c) sum(c!=0)>4),]

mat_colors <- list(group = brewer.pal(9, "Set1"))

nb.cols <- 17
Var1 <- colorRampPalette(brewer.pal(8, "Reds"))(nb.cols)
names(Var1) <- unique(col_annotation$`Cancer type`)
anno_colors <- list(Var1 = Var1)

breaksList = seq(0, 2, by = 0.05)
pheatmap(df_missing_normalized_matrix_frequestnidentified,
         #fontsize_col = 5, 
         fontsize_row = 6,
         # cutree_cols = 17,
         color = colorRampPalette(colors = c("black", "yellow", "darkorange"))(length(breaksList)),
         show_rownames = T,
         show_colnames = F,
         main = "Exosome", 
         cluster_cols = F,
         cluster_rows = T,
         annotation_col = col_annotation)
         # annotation_colors = anno_colors)


```



## MHC Heatmap

```{r echo=TRUE, warning=FALSE}

## let read the required libraries
rm(list = ls())

library(tidyr)
library(dplyr)
library(pheatmap)
library(RColorBrewer)
# library(heatmap.plus)
library(viridis)
library(dendextend)

## let's set the directory
# setwd("~/Box Sync/Missing protein project/")

## read the data first
df_missing_normalized <- read.csv("./Data/MHC_SumTickNormalization.csv", skip = 2,row.names = 1)
df_missing_normalized_all <- read.csv("./Data/MHC_SumTickNormalization.csv", header = F)

## annotation column
col_annotation <- data.frame("Cancer type"= t(df_missing_normalized_all[1,2:dim(df_missing_normalized_all)[2]]))
rownames(col_annotation) <- colnames(df_missing_normalized)
colnames(col_annotation) <- c("Cancer type")

### log2 scaling
df_missing_normalized_matrix <- as.matrix(df_missing_normalized)
df_missing_normalized_matrix[is.na(df_missing_normalized_matrix)] = 0
df_missing_normalized_matrix[df_missing_normalized_matrix!=0] = log10(df_missing_normalized_matrix[df_missing_normalized_matrix!=0])

# ### z score normaliztion instead
# df_missing_normalized_matrix <- as.matrix(df_missing_normalized)
# df_missing_normalized_matrix[is.na(df_missing_normalized_matrix)] = 0
# df_missing_normalized_matrix <- apply(df_missing_normalized_matrix, 2, scale)

### heatmap
pheatmap(df_missing_normalized_matrix,
         #fontsize_col = 5, 
         #fontsize_row = 4,
         show_rownames = F,
         show_colnames = F,
         main = "MHC", 
         cluster_cols = F,
         cluster_rows = F,
         annotation_col = col_annotation)

#### hmmm! well it is so spares, let's focus on those being repeated
df_missing_normalized_matrix_frequestnidentified <- df_missing_normalized_matrix[apply(df_missing_normalized_matrix, 1, function(c) sum(c!=0)>1),]

mat_colors <- list(group = brewer.pal(9, "Set1"))

nb.cols <- 17
Var1 <- colorRampPalette(brewer.pal(8, "Reds"))(nb.cols)
names(Var1) <- unique(col_annotation$`Cancer type`)
anno_colors <- list(Var1 = Var1)

breaksList = seq(0, 2, by = 0.05)
pheatmap(df_missing_normalized_matrix_frequestnidentified,
         #fontsize_col = 5, 
         fontsize_row = 6,
         # cutree_cols = 17,
         color = colorRampPalette(colors = c("black", "yellow", "darkorange"))(length(breaksList)),
         show_rownames = T,
         show_colnames = F,
         main = "MHC", 
         cluster_cols = F,
         cluster_rows = T,
         annotation_col = col_annotation,
         annotation_colors = anno_colors)


```


### Distribution of Samples detected
```{r}

## let read the required libraries
rm(list = ls())
library(readxl)

## let's set the directory
# setwd("~/Box Sync/Missing protein project/")

## read the data first
# df_missing <- read_xlsx("./Data/Table1.xlsx",sheet = "Acc_Coverage")

df_missing$`N Samples Detected (Total 513 Cell Line and Patient Samples)`[df_missing$`N Samples Detected (Total 513 Cell Line and Patient Samples)` >= 50] = 50
hist(df_missing$`N Samples Detected (Total 513 Cell Line and Patient Samples)`, col = "red", breaks = 100, xlab = "Number of Samples", ylab = "Frequency", xlim = c(0,50))
```


### Distribution of PSM
```{r}

## let read the required libraries
rm(list = ls())
library(readxl)

## let's set the directory
# setwd("~/Box Sync/Missing protein project/")

## read the data first
df_missing <- read_xlsx("./Data/Table1.xlsx",sheet = "Table1_data")

df_missing$TotalPep[df_missing$TotalPep >= 100] = 100
hist(df_missing$TotalPep, col = "blue", breaks = 100, xlab = "Peptide Spectra Matching", ylab = "Frequency", xlim = c(0,100))

```


### Venny diagram
```{r}

## let read the required libraries
rm(list = ls())
library(readxl)

## let's set the directory
# setwd("~/Box Sync/Missing protein project/")

## read the data first
df_missing <- read_xlsx("./Data/Table1.xlsx",sheet = "Table1_data")
df_missing_plasma <- read_xlsx("./Data/From Jen_UniquePeptides/UniquePeptides.xlsx")


# Load library
library(VennDiagram)
 
# venn diagram

set1 <- df_missing_plasma$Accession
set2 <- df_missing$Accession[with(df_missing, ExperimentType == "CELLLINE(EXOSOME)")]
set3 <- df_missing$Accession[with(df_missing, ExperimentType == "CELLLINE" | ExperimentType == "PRIMARY CELL")]

myCol <- brewer.pal(3, "Pastel2")

# Chart
venn.diagram(
        x = list(set1, set2, set3),
        category.names = c("Human Plasma" , "Exosome " , "Cancer cell"),
        filename = '#14_venn_diagramm.png',
        output=TRUE,
        
        # Output features
        imagetype="png" ,
        height = 480 , 
        width = 480 , 
        resolution = 300,
        compression = "lzw",
        
        # Circles
        lwd = 2,
        lty = 'blank',
        fill = myCol,
        
        # Numbers
        cex = .5,
        fontface = "bold",
        fontfamily = "sans",
        
        # Set names
        cat.cex = 0.01,
        cat.fontface = "bold",
        cat.default.pos = "outer",
        cat.pos = c(-27, 27, 135),
        cat.dist = c(0.055, 0.055, 0.085),
        cat.fontfamily = "sans",
        rotation = 2
)


### what about cell line compartment
set3 <- df_missing$Accession[with(df_missing, (ExperimentType == "CELLLINE" & Compartment == "TCE"))]
set4 <- df_missing$Accession[with(df_missing, (ExperimentType == "CELLLINE" & Compartment == "Surface"))]
set5 <- df_missing$Accession[with(df_missing, (ExperimentType == "CELLLINE" & Compartment == "Media"))]
set6 <- df_missing$Accession[with(df_missing, (ExperimentType == "CELLLINE" & Compartment == "Nuclear"))]


myCol <- brewer.pal(4, "Pastel2")

# Chart
venn.diagram(
        x = list(set3, set4, set5, set6),
        category.names = c("TCE" , "Surface " , "Media", "Nuclear"),
        filename = '#15_venn_diagramm.png',
        output=TRUE,
        
        # Output features
        imagetype="png" ,
        height = 480 , 
        width = 600 , 
        resolution = 300,
        compression = "lzw",
        
        # Circles
        lwd = 2,
        lty = 'blank',
        fill = myCol,
        
        # Numbers
        cex = .6,
        fontface = "bold",
        fontfamily = "sans",
        
                # Set names
        cat.cex = 0.5,
        cat.fontface = "bold",
        cat.default.pos = "outer",
)


```


### Here, let's focus on correlation between RNA Seq data and proteomic dataset.

```{r}

## let read the required libraries
rm(list = ls())
library(readxl)
library(pROC)
library(tidyr)
library(dplyr)

## let's set the directory
# setwd("~/Box Sync/Missing protein project/")

df_RNA <- read.csv("./Data/Gastric/From Candy/merged_IPAS_RNAseq_counts.csv")
df_RNA <- df_RNA[-c(1:5),]

### since gene name are in ACS form, let's convert it into gene name!
library(EnsDb.Hsapiens.v79)

# 1. Convert from ensembl.gene to gene.symbol
ensembl.genes_split <- lapply(df_RNA$X, function(c) unlist(strsplit(as.character(c), "[.]"))[1])
ensembl.genes_split <- unlist(ensembl.genes_split)
df_RNA$X <- ensembl.genes_split

geneIDs1 <- ensembldb::select(EnsDb.Hsapiens.v79, keys= ensembl.genes_split, keytype = "GENEID", columns = c("SYMBOL","GENEID"))

### merge with the original dataset
df_RNA_withGeneName <- merge(geneIDs1, df_RNA, by.x = "GENEID", by.y = "X")

## Let's normalize the data
df_RNA_withGeneName_Norm <- df_RNA_withGeneName
df_RNA_withGeneName_Norm[,3:11] <- apply(df_RNA_withGeneName_Norm[,3:11],2, function(c) c/sum(c)*10^6)

write.csv(df_RNA_withGeneName_Norm,"./Data/df_RNA_withGeneName_Norm.csv")
### great! I have gene name for each gene Id

#### let's read the protein file.
df_Protein <- read.csv("./Data/Gastric/PatientCompartment=TCE_Normalized_Gastric_withGeneName_Filtered.csv")


### let's merge protein and gene
df_all_Pro <- read_xlsx("./Data/Gastric/RNASeq_withGeneName_Norm_withProteinInfo.xlsx", sheet = "Proteins")
df_all_gene <- read_xlsx("./Data/Gastric/RNASeq_withGeneName_Norm_withProteinInfo.xlsx", sheet = "Genes")

df_all_merge <- merge(df_all_Pro, df_all_gene, by.x = "GeneName", by.y = "SYMBOL")

write.csv(df_all_merge, "./Output/Gene_Protein_Gastric.csv")

# Great! let's do the correlation!
df_all_merge_filter <- df_all_merge[apply(df_all_merge[,3:10], 1, function(c) sum(!is.na(c)))>2,]

df_all_merge_filter <- df_all_merge[apply(df_all_merge[,3:10], 1, function(c) sum(c[!is.na(c)]))>20,]


List_cor <- NA
Name <- NA
for (i in 1:dim(df_all_merge_filter)[1]){
        print(i)
        c <- df_all_merge_filter[i,]
        List_cor[i] <- cor(c[3:10][!is.na(c[3:10])],c[12:19][!is.na(c[3:10])], method = "spearman")
        Name[i] <- df_all_merge_filter$Accession[i]
}

df_Name_cor <- data.frame(Name,List_cor)
### let's compare with totally randome data (let's confirm with Dr. James, Can I have different lenght for random data?)
X <- matrix(sample(50000*8), ncol = 8)
Y <- matrix(sample(50000*8), ncol = 8)

All <- data.frame(X,Y)

List_cor_Randome <- NA
for (i in 1:50000){
        r <- All[i,]
        List_cor_Randome[i] <- cor(t(r[1:8]),t(r[9:16]), method = "spearman")
}

# Filled Density Plot
d <- density(List_cor[!is.na(List_cor)])
plot(d, main="Density of correlation", xlim = c(-1,1))
polygon(d, col="orange", border="black", xlim = c(-1,1))

library(ggplot2)

#Sample data
dat <- data.frame(dens = c(List_cor_Randome, List_cor)
                   , lines = rep(c("randome","z_missing Proteins"),times = c(50000,dim(df_all_merge_filter)[1])))
#Plot.
ggplot(dat, aes(x = dens, fill = lines)) + geom_density(alpha = 0.8) + theme_bw() +scale_fill_manual( values = c("lightgray","lightblue"))


### let's do the heamtmap of identified assecsion!

## read the data first
df_missing_normalized <- read.csv("./Data/CellLineCompartment_TCE_SumTickNormalization.csv",sep = "\t", skip = 2,row.names = 1)
df_missing_normalized_all <- read.csv("./Data/CellLineCompartment_TCE_SumTickNormalization.csv",sep = "\t", header = F)

## annotation column
col_annotation <- data.frame("Cancer type"= t(df_missing_normalized_all[1,2:dim(df_missing_normalized_all)[2]]))

col_annotation$X2 <- as.character(col_annotation$X1)
col_annotation$X2[grep("Breast",col_annotation$X1)] = ("Breast")
col_annotation$X2[grep("Leukemia",col_annotation$X1)] = ("Leukemia")
col_annotation <- data.frame(col_annotation[,-c(1)])
rownames(col_annotation) <- colnames(df_missing_normalized)
colnames(col_annotation) <- c("Simplified Cancer type")


### log2 scaling
df_missing_normalized_matrix <- as.matrix(df_missing_normalized)
df_missing_normalized_matrix[is.na(df_missing_normalized_matrix)] = 0
df_missing_normalized_matrix[df_missing_normalized_matrix!=0] = log10(df_missing_normalized_matrix[df_missing_normalized_matrix!=0])


#### hmmm! well it is so spares, let's focus on those being repeated
df_missing_normalized_matrix_frequestnidentified <- df_missing_normalized_matrix[rownames(df_missing_normalized_matrix) %in% Name,]

# mat_colors <- list(group = brewer.pal(sum(!duplicated(col_annotation$`Simplifeid Cancer type`)), "Set3"))

nb.cols <- sum(!duplicated(col_annotation$`Simplifeid Cancer type`))
Var1 <- colorRampPalette(brewer.pal(8, "Pastel1"))(nb.cols)
names(Var1) <- unique(col_annotation$`Simplifeid Cancer type`)
anno_colors <- list(Var1 = Var1)

breaksList = seq(0, 2, by = 0.05)
pheatmap(df_missing_normalized_matrix_frequestnidentified,
         #fontsize_col = 5, 
         fontsize_row = 8,
         # cutree_rows = 6,
         #color = colorRampPalette(colors = c("black", "yellow", "darkorange"))(length(breaksList)),
         show_rownames = T,
         show_colnames = F,
         main = "Total Cell Extract", 
         cluster_cols = F,
         cluster_rows = T,
         annotation_colors = anno_colors,
         annotation_col = col_annotation
         )


```



## distribution of deteceted samples:

```{r}

## let read the required libraries
rm(list = ls())
library(readxl)
library(pROC)
library(tidyr)
library(dplyr)

## let's set the directory
setwd("~/Box Sync/Missing protein project/")


## let's read the table
df_allproteins <- read_xlsx("./Data/Table1.xlsx", sheet = "Acc_Coverage")

## let's dichotomize the number of samples detected
df_allproteins$dich_smples <- df_allproteins$`N Samples Detected (Total 513 Cell Line and Patient Samples)`
df_allproteins$numbersmpls <- df_allproteins$`N Samples Detected (Total 513 Cell Line and Patient Samples)`


df_allproteins$dich_smples[with(df_allproteins, which(numbersmpls>=0 & numbersmpls <5))] <- "[0-5)"
df_allproteins$dich_smples[with(df_allproteins, which(numbersmpls>=5 & numbersmpls <10))] <- "[5-10)"
df_allproteins$dich_smples[with(df_allproteins, which(numbersmpls>=10 & numbersmpls <25))] <- "[10-25)"
df_allproteins$dich_smples[with(df_allproteins, which(numbersmpls>=25 & numbersmpls <50))] <- "[25-50)"
df_allproteins$dich_smples[with(df_allproteins, which(numbersmpls>=50 & numbersmpls <100))] <- "[50-100)"
df_allproteins$dich_smples[with(df_allproteins, which(numbersmpls>=100 & numbersmpls <200))] <- "[100-200)"
df_allproteins$dich_smples[with(df_allproteins, which(numbersmpls>=200))] <- "200+"

barplot(table(df_allproteins$dich_smples))


### also, let's plot the pie chart
df_IPAout <- read.csv("./Data/IPA_forallmissingproteins.csv")

summary <- data.frame(table(df_IPAout$Type.s.))
summary <- summary[summary$Var1 != "other",]

# 3D Exploded Pie Chart
library(plotrix)
pie3D(summary$Freq,explode=0)

pie3D(summary$Freq, labels = summary$Var1,explode=0)

```


## let's focus on plasma samples right now
```{r}
## let read the required libraries
rm(list = ls())
library(readxl)
library(pROC)
library(tidyr)
library(dplyr)

## let's set the directory
setwd("~/Box Sync/Missing protein project/")

df_plasma <- read_xlsx("./Data/Plasma_AllQuant_missingProteins.xlsx")

Gmean <- log2(df_plasma$GMean)
hist(Gmean[!is.na(Gmean)], breaks = 50, col="red")
hist(Gmean[!is.na(Gmean)], breaks = 50, col="grey")

## how many unique proteins above FC of 2?


df_plasma_hFC <- df_plasma[with(df_plasma, which(GMean>=2)),]
sum(!duplicated(df_plasma_hFC$`protein name`))
```


### okay, Sam suggested new analysis. First, how many unique peptide?
```{r}
## let read the required libraries
rm(list = ls())
library(readxl)
library(pROC)
library(tidyr)
library(dplyr)

## let's set the directory
setwd("~/Box Sync/Missing protein project/")

## let's read the unique peptide

df_uniquepeptides <- read_xlsx("./Data/From Jen_UniquePeptides/UniquePeptides.xlsx")
df_uniquepeptides$modefied <- df_uniquepeptides$`N(UniquePeptides)`
df_uniquepeptides$modefied[df_uniquepeptides$modefied>=25] = 25


hist(df_uniquepeptides$modefied, breaks = 35, col = "lightblue")

```


### what about number of LC-MS versus total numnber of detected proteins?
```{r}
## let read the required libraries
rm(list = ls())
library(readxl)
library(pROC)
library(tidyr)
library(dplyr)

## let's set the directory
setwd("~/Box Sync/Missing protein project/")

## now, let's read the file
Fractions <- read.csv("./Data/Fractions.csv")

Tablerepeat <- data.frame(table(Fractions$File))
Tablerepeat$Runnumber <- 1:dim(Tablerepeat)[1]
write.csv(Tablerepeat,"./Output/NumberofRun_identifiedFeatures.csv")

```


### Let's look at the plasma! only those that are not expressed in heathly controls

```{r}

## let read the required libraries
rm(list = ls())
library(readxl)
library(pROC)
# library(tidyr)
library(dplyr)
library(tidyr)
library(dplyr)
library(pheatmap)
library(RColorBrewer)
# library(heatmap.plus)
library(viridis)
library(dendextend)


## let's set the directory
# setwd("~/Box Sync/Missing protein project/")

## let's read the file
df_plasma <- read_xlsx("./Data/From Jen_UniquePeptides/Plasma_AllQuant_missingProteins_EhsanVersion.xlsx", sheet = "Plasma_AllQuant_peptides")

df_plasma_short <- df_plasma[,c(2,3,6,10,12,13)]


df_plasma_short_compressed <- df_plasma_short %>% group_by(Acc,gene,IPI,Assay) %>% summarise_each(sum)

df_plasma_short_compressed <- df_plasma_short_compressed[with(df_plasma_short_compressed, which(Q3L < 1)),]

df_plasma_sampleinfo <- read_xlsx("./Data/From Jen_UniquePeptides/Plasma_AllQuant_missingProteins_EhsanVersion.xlsx", sheet = "SupplementExperiment")


df_plasma_short_compressed_wSampleinfo <- merge(df_plasma_short_compressed,df_plasma_sampleinfo, by.x = "Assay", by.y="IPAS")


df_plasma_short_compressed_wSampleinfo_selecteddisease <- df_plasma_short_compressed_wSampleinfo[with(df_plasma_short_compressed_wSampleinfo, which(Disease == "Breast" | Disease == "Colon" | Disease == "Lung" | Disease == "Pancreatic")),]

## great! Let's overlap them with tissue expression

dfnormalTissue <- read.table(file = '~/Desktop/big files/rna_tissue_consensus.tsv', sep = '\t', header = TRUE)

library(reshape2)
dfnormalTissue_matrix <- acast(dfnormalTissue, dfnormalTissue$Gene.name~dfnormalTissue$Tissue,value.var = "NX", fun.aggregate = sum)

dfnormalTissue_matrix_backup <- dfnormalTissue_matrix

dfnormalTissue_matrix <- apply(dfnormalTissue_matrix, 1, function(c) c/sum(c)*10)
dfnormalTissue_matrix_t <- data.frame(t(dfnormalTissue_matrix))
dfnormalTissue_matrix_t$gene <- rownames(dfnormalTissue_matrix_t)
### merge with plasma samples



df_merge_selected_tissue <- merge(df_plasma_short_compressed_wSampleinfo_selecteddisease, dfnormalTissue_matrix_t, by.x = "gene", by.y = "gene", all.x = TRUE)

write.csv(df_merge_selected_tissue, "./Output/onlycasesinPlasma_overlapwithTissue.csv")

## let's sort it by disease
df_merge_selected_tissue <- df_merge_selected_tissue[order(df_merge_selected_tissue$Disease),]
## now, let's plot the heatmap
df_heatmapplasma <- as.matrix(df_merge_selected_tissue[,c(24:dim(df_merge_selected_tissue)[2])])
df_heatmapplasma[is.na(df_heatmapplasma)] <- 0
rownames(df_heatmapplasma) <- paste0(df_merge_selected_tissue$Assay," ",df_merge_selected_tissue$Acc)

## rowannotation
df_heatmapplasma_rowannot <- as.matrix(df_merge_selected_tissue[,"Disease"])
rownames(df_heatmapplasma_rowannot) <- rownames(df_heatmapplasma)
df_heatmapplasma_rowannot <- as.data.frame(df_heatmapplasma_rowannot)
colnames(df_heatmapplasma_rowannot) <- "Cancer type"

breaksList = seq(0, 10, by = 0.1)
pheatmap(as.data.frame(df_heatmapplasma),
         fontsize_col = 7, 
         fontsize_row = 3,
         # cutree_rows = 6,
         color = colorRampPalette(colors = c("lightblue", "white", "red"))(length(breaksList)),
         show_rownames = T,
         show_colnames = T,
         main = "Normal Tissue expression", 
         cluster_cols = F,
         cluster_rows = F,
         annotation_row = df_heatmapplasma_rowannot
         )

```

