---
title: "analysis_combined"
output: html_document
date: "2022-09-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(DEqMS)
library(tidyverse)
library(pheatmap)
```


# Reading data
```{r}
SpC_TCE <- read.delim("data/CellLineCompartment_TCE_SumTickNormalization.tsv", sep = '\t', skip=2, row.names = 'Accession')
SpC_surface <- read.delim("data/CellLineCompartment_Surface_SumTickNormalization.tsv",  sep = '\t', skip=2, row.names = 'Accession')
SpC_exosome <- read.csv("data/CellLineCompartment_Exosome_SumTickNormalization.csv", skip=2, row.names = 'Accession')
SpC_media <- read.csv("data/CellLineCompartment_Media_SumTickNormalization.csv", skip=2, row.names = 'Accession')
SpC_nuclear <- read.delim("data/CellLineCompartment_Nuclear_SumTickNormalization.csv", sep = '\t', skip=2, row.names = 'Accession')
SpCs <- list(TCE=SpC_TCE, surface=SpC_surface, exosome=SpC_exosome, media=SpC_media, nuclear=SpC_nuclear)
```



```{r}
row_not_all_missing <- function(X) {
  # X[X == 0] <-  NA
  is_missing <- is.na(X)
  row_keep <- (rowSums(is_missing) != ncol(X))
  return(rownames(X)[row_keep])
}

MPs_in_components <- lapply(SpCs, row_not_all_missing)
```


## create column annotation
```{r}
extract_annotation <- function(SpC) {
  disease_fulltype <- str_extract(colnames(SpC), '.+(?=\\.IP)')
  disease <- disease_fulltype
  disease[str_detect(disease_fulltype, 'Breast')] <- 'Breast'
  disease[str_detect(disease_fulltype, 'Leukemia')] <- 'Leukemia'
  IPAS_annotation <- data.frame(disease = disease, row.names = colnames(SpC))
}

annotations <- lapply(SpCs, extract_annotation)
```

## read-in newest missing protein list
```{r}
missing_protein_df <- readxl::read_xlsx('data/PE2-5.xlsx')
missing_protein_df <- missing_protein_df[-c(1:12), ]
# head(missing_protein_df)
```



# analysis
integrate information to create SpC_list objects in a list
```{r}
SpC_lists <- mapply(SpC_List, SpCs, annotations, MoreArgs = list(NA_substitution = NULL, proteins_filter = missing_protein_df$Accession), SIMPLIFY = FALSE)
```

## check cancer type distribution
The minority group in different component dataset is different, if I collapse minorities in each of the
dataset into "other", the group "other" is not comparable between dataset.
So, I prefer just remove minority groups in each components dataset by a rule of minimum 3

TCE: neuroblastoma (2) lymphoblast(2) 
SurfaceL neuroblastoma (2) 
exsome: NSCLC (1)
media: Leukemia (1) neuroblastoma (2)
nuclear: Ovarian (2)
```{r}
basic_describe(SpC_lists$TCE, plot_boxplot = FALSE, piechart_main = 'Total cell extraction (TCE)')
par(mar=c(0.5, 0.5, 1, 1.6), mfrow=c(2, 2))
mapply(function(x, title) {basic_describe(x, plot_boxplot=FALSE, piechart_main = title)}, x=SpC_lists[-1], title=names(SpC_lists[-1]))
```


```{r}
par(mar=c(1, 1, 1, 1.6), mfrow=c(2, 3))
mapply(function(x, title) {basic_describe(x, plot_piechart=FALSE, boxplot_main = title)}, x=SpC_lists, title=names(SpC_lists))
```



Remove the minority cancer type from each dataset
```{r}
extract_sample_id <- function(x, disease) {
  rownames(x$annotation[x$annotation$disease %in% disease, ,drop=FALSE])
}

minority_diseases <- list(TCE=c('Neuroblastoma', 'Lymphoblast'),
                          surface=c('Neuroblastoma'),
                          exosome=c('NSCLC'),
                          media=c('Neuroblastoma', 'Leukemia'),
                          nuclear=c('Ovarian'))
replicates_for_removing <- mapply(extract_sample_id, SpC_lists, minority_diseases, SIMPLIFY = FALSE)
replicates_for_removing$TCE <- c(replicates_for_removing$TCE, 'Prostate.IP0872_1701.H660')
```


```{r}
SpC_lists <- mapply(SpC_List, SpCs, annotations, replicates_remove=replicates_for_removing, MoreArgs = list(NA_substitution=NULL, proteins_filter=missing_protein_df$Accession), SIMPLIFY = FALSE)
```


```{r}
par(omr=c(0.5, 0.5, 0.5, 1), mar=c(1, 4, 1, 4), mfrow=c(1, 2))
basic_describe(SpC_lists$TCE, piechart_main = 'Total cell extraction (TCE)')
par(mar=c(0.5, 0.5, 1, 1.6), mfrow=c(2, 2))
mapply(function(x, title) {basic_describe(x, pie_boxplot=FALSE, boxplot_main = title, piechart_main = title)}, x=SpC_lists[-1], title=names(SpC_lists[-1]))
```


# Section 1: Missing proteins localization


# Data pre-filter
Since the data is normalized not raw data, not use a pre-filter here
<!-- remove proteins that is not confidentially detected -->
```{r}
prefilter <- function(x, count_threshold) {
  N <- min(table(x$annotation$disease))
  keep <- rowSums(x$matrix > 1, na.rm = T) >= N
  return(rownames(x$matrix[keep,]))
}


# get MPs 
MPs <- lapply(SpC_lists, prefilter)
sapply(MPs, length)


# compare with number of MPs before pre-filtering 
MPs_before_filter <- lapply(SpC_lists, function(x) {return(rownames(x$matrix))})
sapply(MPs_before_filter, length)
```


```{r}
library(VennDiagram)
library(RColorBrewer)
myCol <- brewer.pal(4, "Pastel2")

venn.diagram(
        x = MPs[-1],
        category.names = names(MPs[-1]),
        filename = 'output/MPs_after_filter.png',
        output=TRUE,
        
        main='Missing proteins after pre-filtering',
        main.cex = 0.5,
        
        # Output features
        imagetype="png" ,
        height = 480 , 
        width = 600 , 
        resolution = 300,
        compression = "lzw",
        
        # Circles
        lwd = 2,
        lty = 'blank',
        fill = myCol,
        
        # Numbers
        cex = .6,
        fontface = "bold",
        fontfamily = "sans",
        
                # Set names
        cat.cex = 0.5,
        cat.fontface = "bold",
        cat.default.pos = "outer"
)
```

```{r}
myCol <- brewer.pal(4, "Pastel2")

venn.diagram(
        x = MPs_in_components[-3],
        category.names = names(MPs_in_components[-3]),
        filename = 'output/MPs_before_filter.png',
        output=TRUE,
        
        main='Missing proteins before pre-filtering',
        main.cex = 0.5,
        
        # Output features
        imagetype="png" ,
        height = 480 , 
        width = 600 , 
        resolution = 300,
        compression = "lzw",
        
        # Circles
        lwd = 2,
        lty = 'blank',
        fill = myCol,
        
        # Numbers
        cex = .6,
        fontface = "bold",
        fontfamily = "sans",
        
                # Set names
        cat.cex = 0.5,
        cat.fontface = "bold",
        cat.default.pos = "outer",
)
```


# Gene Oncology analysis
Not using GSEA because the it require fold-change of DEGs and the we are only infer
a DEGs based on DEPs so the fold-change of genes is not available
The reason for not run pathway analysis is the same, current data can only used for a
crude analysis
```{r}
missing_protein_df
```


```{r}
df <- missing_protein_df %>% dplyr::select(Accession, `gene name(s)`) 
df <- base::data.frame(gene_name = df$`gene name(s)`, row.names = df$Accession)
```


```{r}
library(org.Hs.eg.db)

mapping_to_entrez <- function(proteins) {
  gene_names <- df[proteins,]
  entrez_id <- AnnotationDbi::select(org.Hs.eg.db, gene_names, keytype = "SYMBOL", columns=c("ENTREZID", "UNIPROT"))$ENTREZID
  return(entrez_id)
}

# uniprots <- MPs$surface
# gene_names <- df[uniprots,]
```


```{r}
MPs_entrez <- lapply(MPs, mapping_to_entrez)
GOs <- lapply(MPs_entrez, goana, species = 'Hs')
```


```{r}
top_GOs_BP <- lapply(GOs, topGO, number = 50, ontology='BP', truncate=30)
top_GOs_CC <- lapply(GOs, topGO, number = 50, ontology='CC', truncate=30)
top_GOs_MF <- lapply(GOs, topGO, number = 50, ontology='MF', truncate=30)
```




```{r}
par(mfrow=c(2, 2))
plot_venn_diagram(top_GOs_BP, main='Biology pathway top 50')
plot_venn_diagram(top_GOs_CC, main='Cellular component top 50')
plot_venn_diagram(top_GOs_MF, main='Molecular function top 50')
```


```{r}
par(mfrow=c(3, 2))
plts <- lapply(top_GOs_BP, bubble_plot)
plts
```


```{r}
par(mfrow=c(3, 2))
plts <- lapply(top_GOs_CC, bubble_plot)
plts
```

```{r}
par(mfrow=c(3, 2))
plts <- lapply(top_GOs_MF, bubble_plot)
plts
```





# Section 1.2 cross-validation with RNA-seq data (correlation)
```{r}
read.csv("data/CellLineCompartment_TCE_SumTickNormalization.tsv", sep = '\t', skip=2, row.names = 'Accession')
```










# section 2: missing protein by cancer type
```{r}
res_lst <- one_vs_all_test(SpC_lists$TCE, method = 'Wilcoxon', logFC = 0, PAjusted = 1)
```


