---
title: "analysis_combined"
output: html_document
date: "2022-09-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(DEqMS)
library(tidyverse)
library(pheatmap)
```


# import data
```{r}
nasf_TCE <- read.csv('data/nasf/nasf_TCE_missing_protein.csv', row.names = 1)
nasf_media <- read.csv('data/nasf/nasf_media_missing_protein.csv', row.names = 1)
nasf_surface <- read.csv('data/nasf/nasf_surface_missing_protein.csv', row.names = 1)
nasf_nuclear <- read.csv('data/nasf/nasf_nuclear_missing_protein.csv', row.names = 1)
nasfs <- list(TCE=nasf_TCE,  surface=nasf_surface, media=nasf_media, nuclear=nasf_nuclear)
```

# import annotation
```{r}
IPAS_annotation <- read_csv('data/support data/IPAS_annotation.csv')
```
annotations
```{r}
create_annoataion <- function(data, annotation_table) {
  annotation_tbl <- annotation_table %>% select(ipas, disease) %>% filter(ipas %in% colnames(data))
  annotation_df <- data.frame(disease=annotation_tbl$disease)
  rownames(annotation_df) <- annotation_tbl$ipas
  return(annotation_df)
}

create_annoataion(nasfs$nuclear, annotation_table = IPAS_annotation)

annotations <- lapply(nasfs, create_annoataion, annotation_table=IPAS_annotation)
```


integrate information to create SpC_list objects in a list
```{r}
SpC_lists <- mapply(SpC_List, nasfs, annotations, SIMPLIFY = FALSE)
```

## check cancer type distribution
The minority group in different component dataset is different, if I collapse minorities in each of the
dataset into "other", the group "other" is not comparable between dataset.
So, I prefer just remove minority groups in each components dataset by a rule of minimum 3

TCE: neuroblastoma (2) lymphoblast(2) 
SurfaceL neuroblastoma (2) 
exsome: NSCLC (1)
media: neuroblastoma (2)
nuclear:  ovarian (1)
```{r}
basic_describe(SpC_lists$TCE, plot_boxplot = FALSE, piechart_main = 'Total cell extraction (TCE)')
par(mar=c(0.5, 0.5, 1, 1.6), mfrow=c(2, 2))
mapply(function(x, title) {basic_describe(x, plot_boxplot=FALSE, piechart_main = title)}, x=SpC_lists, title=names(SpC_lists))
```


```{r}
par(mar=c(1, 1, 1, 1.6), mfrow=c(2, 3))
mapply(function(x, title) {basic_describe(x, plot_piechart=FALSE, boxplot_main = title)}, x=SpC_lists, title=names(SpC_lists))
```



Remove the minority cancer type from each dataset
```{r}
extract_sample_id <- function(x, disease) {
  rownames(x$annotation[x$annotation$disease %in% disease, ,drop=FALSE])
}

minority_diseases <- list(TCE=c('Neuroblastoma', 'Lymphoblast'),
                          surface=c('Neuroblastoma'),
                          media=c('Neuroblastoma', 'Leukemia'),
                          nuclear=c('Ovarian'))

replicates_for_removing <- mapply(extract_sample_id, SpC_lists, minority_diseases, SIMPLIFY = FALSE)
```


```{r}
SpC_lists <- mapply(SpC_List, nasfs, annotations, replicates_remove=replicates_for_removing, SIMPLIFY = FALSE)
```


```{r}
par(mar=c(0.5, 0.5, 1, 1.6), mfrow=c(2, 2))
mapply(function(x, title) {basic_describe(x, pie_boxplot=FALSE, boxplot_main = title, piechart_main = title)}, x=SpC_lists, title=names(SpC_lists))
```


# Section 1: Missing proteins localization


# Data pre-filter
Since the data is normalized not raw data, not use a pre-filter here
<!-- remove proteins that is not confidentially detected -->
```{r}
prefilter <- function(x, count_threshold) {
  N <- min(table(x$annotation$disease))
  keep <- rowSums(x$matrix > 0, na.rm = T) >= N
  return(rownames(x$matrix[keep,]))
}

# get MPs 
MPs_after_prefilter <- lapply(SpC_lists, prefilter)
sapply(MPs_after_prefilter, length)


# compare with number of MPs before pre-filtering 
MPs_before_prefilter <- lapply(SpC_lists, function(x) {return(rownames(x$matrix))})
sapply(MPs_before_prefilter, length)

# export the data
save(MPs_after_prefilter, file='cache/MPs_after_prefilter.rda')
```


```{r}
plot_venn_diagram(MPs_before_prefilter, main = 'Missing proteins before pre-filtering')
```

```{r}
plot_venn_diagram(MPs_after_prefilter, main = 'Missing proteins after pre-filtering')
```



# Gene Oncology analysis
Not using GSEA because the it require fold-change of DEGs and the we are only infer
a DEGs based on DEPs so the fold-change of genes is not available
The reason for not run pathway analysis is the same, current data can only used for a
crude analysis
```{r}
# missing_protein_df <- readxl::read_xlsx('data/support data/PE2-5.xlsx')
# missing_protein_df <- missing_protein_df[-c(1:12), ]
load('cache/missing_protein.rda')
df <- missing_protein_df %>% dplyr::select(Accession, `gene name(s)`) 
df <- base::data.frame(gene_name = df$`gene name(s)`, row.names = df$Accession)
```


```{r}
mapping_to_entrez <- function(proteins) {
  gene_names <- df[proteins,]
  entrez_id <- AnnotationDbi::select(org.Hs.eg.db::org.Hs.eg.db, gene_names, keytype = "SYMBOL", columns=c("ENTREZID", "UNIPROT"))$ENTREZID
  return(entrez_id)
}
```


```{r}
MPs_entrez <- lapply(MPs_after_prefilter, mapping_to_entrez)
GOs <- lapply(MPs_entrez, goana, species = 'Hs')
```
```{r}
topGO_plus <- function(results, ontology, number = Inf) {
  sorted <- topGO(results, ontology = ontology, number = number)
  return(sorted[sorted$P.DE < 0.05, ])
}
```


```{r}
top_GOs_BP <- lapply(GOs, topGO_plus, ontology='BP')
top_GOs_CC <- lapply(GOs, topGO_plus, ontology='CC')
top_GOs_MF <- lapply(GOs, topGO_plus, ontology='MF')
```




```{r}
par(mfrow=c(2, 2))
plot_venn_diagram_top_Go(top_GOs_BP, main='Top Biology pathways with P < 0.05')
plot_venn_diagram_top_Go(top_GOs_CC, main='Top Cellular component with P < 0.05')
plot_venn_diagram_top_Go(top_GOs_MF, main='Top Molecular function with P < 0.05')
```

## export GO analysis
```{r}
write_csv(top_GOs_BP$TCE, file = 'data/GO analysis/significant_BP_TCE.csv')
```


```{r}
par(mfrow=c(3, 2))
plts <- lapply(top_GOs_BP, bubble_plot)
plts
```


```{r}
par(mfrow=c(3, 2))
plts <- lapply(top_GOs_CC, bubble_plot)
plts
```

```{r}
par(mfrow=c(3, 2))
plts <- lapply(top_GOs_MF, bubble_plot)
plts
```


# cancer DEMPs
```{r}
TCE_SpC <- read.csv('data/cell line raw data/tce_raw_count.csv', row.names = 1)
# annotation <- read_tsv("data/CellLineCompartment_TCE_SumTickNormalization.tsv", n_max=3, col_select = c(-1), col_names = F, show_col_types = FALSE)
```



```{r}

```


```{r}
y <- SpC_List(TCE_SpC, TCE_annotation, NA_substitution = 0, proteins_filter = MPs_after_prefilter$TCE, replicates_remove = replicates_for_removing$TCE)
```

```{r}
par(mfrow=c(1, 2))
basic_describe(y)
```




perform test
```{r}
res_lst <- one_vs_all_test(y, logFC = 1, PAjusted = 0.05)
```


res_lst
```{r}
unique_DEPs_TCE <- unique(unlist(lapply(res_lst, rownames)))
# paste0(unique_DEPs, collapse = '\n')
```



# Heatmap for the entire data matrix
```{r}
heatmap_dat$matrix <- SpC_lists$TCE$matrix[, rownames(SpC_lists$TCE$annotation)]
heatmap_dat$annotation <- SpC_lists$TCE$annotation
heatmap_dat$matrix <- log2(heatmap_dat$matrix + 1)
plot_heat_map(heatmap_dat, row_keep = unique_DEPs_TCE, show_rownames = F, main='Differentially expressed missing proteins in TCE')
```





   Breast         Colon       Gastric        Glioma      Leukemia     LungAdeno   Lymphoblast      Melanoma Neuroblastoma       Ovarian    Pancreatic 
      39           10             9            10            14            60             2             5             2            14            18 
     Prostate          SCLC 
            5            27 
```{r}
unique_proteins_subsets <- lapply(res_lst, rownames)
# check number of missing proteins in each cancer type
# lapply(unique_proteins_subsets, length)

venn_dat <- unique_proteins_subsets[c('SCLC', 'Breast', 'Pancreatic', 'LungAdeno')]

plot_venn_diagram(venn_dat, main = 'Differentially expressed missing proteins in four cancer types', main.cex=1.5)
```


PCA
```{r}
library(ggfortify)
pca_res <- prcomp(t(SpC_lists$TCE$matrix[unique_DEPs_TCE, ]), center = TRUE, scale. = TRUE)
autoplot(pca_res, data = SpC_lists$TCE$annotation, colour = 'disease')
# TCE_SpC[unique_DEPs_TCE, ]
```
t-SNE
```{r}
set.seed(142)

# col_filter <- rownames(SpC_lists$TCE$annotation)[!SpC_lists$TCE$annotation$disease %in% c('Melanoma', 'Prostate')]
# tsne_dat <- SpC_lists$TCE$matrix[unique_DEPs_TCE, col_filter]
# tsne_annotation <- SpC_lists$TCE$annotation[col_filter, , drop=FALSE]
# # align with same order
# tsne_dat <- tsne_dat[, rownames(tsne_annotation)]
# tsne_dat <- t(tsne_dat)

tsne_dat <- SpC_lists$TCE$matrix[MPs_after_prefilter$TCE, ]
tsne_annotation <- SpC_lists$TCE$annotation
tsne_dat <- tsne_dat[, rownames(tsne_annotation)]
tsne_dat <- t(tsne_dat)

tSNE_fit <- tsne_dat %>% 
  Rtsne::Rtsne(check_duplicates=FALSE)

y_meta <- tsne_annotation %>%
  select(disease) %>% 
  mutate(ID=row_number())

tSNE_df <- tSNE_fit$Y %>% 
  as.data.frame() %>%
  rename(tSNE1="V1",
         tSNE2="V2") %>%
  mutate(ID=row_number())

tSNE_df <- tSNE_df %>%
  inner_join(y_meta , by="ID")



tSNE_df %>%
  ggplot(aes(x = tSNE1, 
             y = tSNE2,
             color = disease
              
            ))+
  geom_point(size = 3)+
  theme(legend.position="bottom")

```


```{r}
missing_DEPs_info <- missing_protein_df %>% filter(Accession %in% unique_DEPs)
```



```{r}
pie(table(missing_DEPs_info$PE))
```







