---
title: "data_preparation"
output:
  pdf_document: default
  html_document: default
date: "2022-08-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(edgeR)
library(pheatmap)
library(RColorBrewer)
library(EnhancedVolcano)
```


```{r}
SpC <- read_csv('data/ALL_Protein_MS2.csv', skip = 3, show_col_types=FALSE)
head(SpC)
```

# create annotation table
```{r}
IPAS_annotation <- read_csv('data/ALL_Protein_MS2.csv', n_max = 4, col_select = -1, col_names = FALSE, show_col_types = FALSE)
IPAS_annotation <- IPAS_annotation %>% t()
col_name <- c('cell_line', 'disease', 'subtype', 'IPAS')
IPAS_annotation <- IPAS_annotation[-1, ] %>% as_tibble()
colnames(IPAS_annotation) <- col_name
head(IPAS_annotation)
```




# accession map to gene dictionary
```{r}
dat_for_mapping_acc_to_gene <- SpC %>% 
  select(Gene, Accession) 
  # only use the primary accession (? will have duplicate row)

acc_to_gene_dict <- dat_for_mapping_acc_to_gene %>% distinct()
```

# collapse cancer category
```{r}
IPAS_annotation <- IPAS_annotation %>% mutate(disease = if_else(disease %in% c('Melanoma', 'Lymphoblast', 'Neuroblastoma', 'Prostate'), 'Other', disease))
```


# Missing protein
```{r}
missing_protein_df <- readxl::read_xlsx('data/PE2-5.xlsx')
missing_protein_df <- missing_protein_df[-c(1:12), ]
head(missing_protein_df)
```

```{r}

```





# ============================================================================
# ============================================================================
```{r}
library(tidyverse)
library(edgeR)
library(pheatmap)
library(RColorBrewer)
library(EnhancedVolcano)
```


# create corresponding count matrix
```{r}
SpC_matrix <- as.matrix(SpC[-c(1, 2)])
rownames(SpC_matrix) <-  pull(SpC[2])


SpC_matrix[is.na(SpC_matrix)] <-  0
```


# Exploration


# Missing data handling 

## missing count

The number of proteins is very high in the merged dataset, could it is because
we use a large target database for searching matched proteins or because each group
of sample has some unique proteins and merged them together create a lot of zero
in the other group?
```{r}
# none of proteins has valid value in all replicates 
SpC %>% summarise(num_of_proteins = n())
SpC %>% drop_na() %>% summarise(num_of_proteins = n())
```

## simple imputation: replace NA as 0
```{r}
# SpC <- SpC %>% mutate_all(~replace(., is.na(.), 0))
SpC_matrix[is.na(SpC_matrix)] <- 0
```


## remove replicates that library size is 0
```{r}
zero_library_filter <- apply(SpC_matrix, 2, sum) != 0
SpC_matrix <- SpC_matrix[, zero_library_filter]
annotation <- IPAS_annotation[zero_library_filter, ]
```



## pre-filter
https://support.bioconductor.org/p/64484/
filter proteins that has at least 5 counts in at least N group
N is the size of the smallest group
However, using count-based model should be robust against low count issue
https://www.mathworks.com/help/stats/negative-binomial-distribution.html

```{r}
N <- min(table(IPAS_annotation$disease))
keep <- rowSums(SpC_matrix > 5) >= N
SpC_matrix_dense <- SpC_matrix[keep,]

# SpC_matrix_dense <- SpC_matrix[apply(SpC_matrix, 1, function(c) sum(c!=0) >= 100), ]
```

recheck overall distribution
```{r}
ggplot(data = data.frame(x=c(SpC_matrix_dense)), aes(x=x)) + 
  geom_histogram(aes(y = ..density..),
                 colour = 1, fill = "white") +
  geom_density(lwd = 1, colour = 4,
               fill = 4, alpha = 0.25)
```

# tag-wise distribution
the majority of proteins should be non-differentially expressed so they will follow a NB distribution
there is no evidence of bimodal issue 

! after find high-confident DEPs, come back to check its distribution (how it violate NB distribution)
```{r}
set.seed(1010)
sample_filter <- sample(1:nrow(SpC_matrix_dense), 100)
density_plot_dat <- t(SpC_matrix_dense[sample_filter, ]) %>% as_tibble() %>% pivot_longer(cols = everything(), names_to = 'Accession', values_to = 'count') 
ggplot(data = density_plot_dat, aes(x=count, color=Accession)) + 
  geom_density(alpha = 0.2) +
  scale_x_continuous(breaks=0:3, limits=c(0, 3)) +
  theme(legend.position="none")
```




# sample(replicate)-wise count distribution 
The distribution they follows isn't important. What is important is whether they 
are similar to each other. Should remove outlier before normalization otherwise
it will bias normalization for the other samples.
```{r}
density_plot_dat <- SpC_matrix_dense %>% as_tibble() %>% pivot_longer(cols = everything(), names_to = 'IPAS', values_to = 'count')
ggplot(data = density_plot_dat, aes(x=log(count+1), fill=IPAS)) + 
  geom_density(alpha = 0.3) +
  scale_x_continuous(breaks=0:5, limits=c(0, 5)) +
  theme(legend.position="none")
```


# create DGEList data class

# need find another way to include gene annotation in the DEGList object
```{r}
# dict <-  deframe(acc_to_gene_dict[, c(2,1)])
# rownames(SpC_matrix_dense) <- dict[rownames(SpC_matrix_dense)]
# genename_filter  <- !is.na(rownames(SpC_matrix_dense))
# SpC_matrix_dense <- SpC_matrix_dense[!is.na(rownames(SpC_matrix_dense)), ]

group <- factor(annotation$disease)
y <- DGEList(SpC_matrix_dense, group = group)
```


<!-- # Filtering -->
<!-- Should I filter protein with too less expression? -->
<!-- prefer not to filter now cause missing proteins are too sparse -->



# normalization
TMM normalization is used.
TMM normalization will adjust library size to ensure that the majority of proteins were not 
differentially expressed.
```{r}
y <- calcNormFactors(y)
```

 

# edgeR estimate overdispersion parameters
Breast is used as control group
```{r}
design <- model.matrix(~group)
colnames(design) <- str_replace(colnames(design), 'group', '')
y <- estimateDisp(y, design = design, robust=TRUE)

fit <- glmQLFit(y, design, robust = TRUE)
```


# ANOVA test
```{r}
ANOVA_test <- glmQLFTest(fit, coef=colnames(design)[-1])
```

# pair-wise test
```{r}
Contrast_Colon_breast_test <- glmQLFTest(fit, coef=c('Colon'))
```




# post-filter
https://pubmed.ncbi.nlm.nih.gov/23770383/
refer to An effect size filter improves the reproducibility in spectral counting-based comparative proteomics


```{r}
ANOVA_DEPs <- topTags(ANOVA_test, Inf, p.value = 0.05)$table
ANOVA_missing_DEPs <- ANOVA_DEPs[rownames(ANOVA_DEPs) %in% missing_protein_df$Accession, ]
mean(rownames(ANOVA_missing_DEPs) %in% unique_DEPs)
```


```{r}
DEPs <- topTags(Contrast_Colon_breast_test, Inf, p.value = 0.05)
```

```{r}
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# 
# BiocManager::install("GO.db")

# goana(test)
```


```{r}
# plotMDS(y, labels = group, col=group)
```

```{r}
plotBCV(y)
```

```{r}
plotQLDisp(fit)
```


```{r}
plotMD(Contrast_Colon_breast_test)
abline(h=c(-1, 1), col="blue")
```


# MDS (PCA)
the first two-dimension only explain 28% percent of variation
MDS plot may not be very useful for clustering.
```{r}
MDS_plot <- plotMDS(y, top=10000, labels = IPAS_annotation$disease)
```


```{r}
ggplot(data=tibble(x=MDS_plot$x, y=MDS_plot$y, type=as.factor(annotation$disease)), aes(x=x, y=y, color=type)) +
  geom_point()
```



# Visulation

## Heatmap for DE proteins in ANOVA test
```{r}
# heatmap_dat <- cpm(y, normalized.lib.sizes = TRUE, log=TRUE, prior.count = 2)
```

```{r}
log_SpC <- log(SpC_matrix_dense[rownames(ANOVA_DEPs$table), ] + 1)
# annotation_col is a dataframe with rownames corresponding to the colnames in the feeded matrix
pheatmap(log_SpC,
         #fontsize_col = 5, 
         #fontsize_row = 4,
         show_rownames = F,
         show_colnames = F,
         main = "TCE", 
         cluster_cols = T,
         cluster_rows = T,
         annotation= data.frame(annotation$Disease, row.names=annotation$IPAS))
```

```{r}
  # if (!requireNamespace('BiocManager', quietly = TRUE))
  #   install.packages('BiocManager')
  # 
  # BiocManager::install('EnhancedVolcano')


 EnhancedVolcano(Contrast_Colon_breast_test$table,
    lab = rownames(Contrast_Colon_breast_test$table),
    x = 'logFC',
    y = 'PValue')
```

# Venn diagram
```{r}

```




# GO ontology and pathway analysis
```{r}
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# 
# BiocManager::install("GO.db")
# BiocManager::install("org.Hs.eg.db")


go <- goana(Contrast_Colon_breast_test, species='Hs')
topGO(go, sort='up')
```



# KEGG
```{r}
kegga(fit)
```


# Missing proteins
```{r}
missing_DEPs <- ANOVA_DEPs$table %>% as_tibble(rownames = 'Accession') %>% filter(Accession %in% missing_protein_df$Accession)
```






```{r}

```




# other processing
1. collapse cancer type category
2. fold-change threshold
3. need corresponding gene name for GO and KEGG analysis






# concerns:
imputation missing value with 0 may not be correct
