---
title: "EdgeR"
output: html_document
date: "2022-08-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# 
# BiocManager::install("edgeR")
```


```{r}
library(tidyverse)
library(edgeR)
library(pheatmap)
library(RColorBrewer)
library(EnhancedVolcano)
```



# Reading data
```{r}
df_missing_normalized <- read_csv("data/raw_spectral_count_all.csv", show_col_types = FALSE)
# df_missing_normalized <- read_tsv("data/CellLineCompartment_TCE_SumTickNormalization.tsv", skip=2)
annotation <- read_tsv("data/CellLineCompartment_TCE_SumTickNormalization.tsv", n_max=3, col_select = c(-1), col_names = F)
```

## create column annotation
```{r}
## annotation column
annotation <- as_tibble(t(annotation))
colnames(annotation) <- c('simplified_cancer_type', 'IP', 'full_name')
annotation <- annotation %>% mutate(simplified_cancer_type = if_else(str_detect(simplified_cancer_type, 'Breast'), 'Breast', 
                                                           if_else(str_detect(simplified_cancer_type, 'Leukemia'), 'Leukemia', simplified_cancer_type)))
```


```{r}
# extract IPAS information
 # paste0(annotation$IP, collapse = ',')
```


# create corresponding count matrix
```{r}
df_missing_normalized_matrix <- as.matrix(df_missing_normalized[-1])
rownames(df_missing_normalized_matrix) <-  pull(df_missing_normalized[1])
# df_missing_normalized_matrix[is.na(df_missing_normalized_matrix)] <-  0
# df_missing_normalized_matrix <- log2(df_missing_normalized_matrix + 1)
```


# Exploration

#check underlying assumption for NB model
```{r}
pseudo_counts <- log(df_missing_normalized_matrix + 1)
ggplot(data = data.frame(x=c(as.matrix(pseudo_counts))), aes(x=x)) + 
  geom_histogram(aes(y = ..density..),
                 colour = 1, fill = "white") +
  geom_density(lwd = 1, colour = 4,
               fill = 4, alpha = 0.25)
```


# use to check normalization status
```{r}
boxplot(pseudo_counts[, 1:10])
```



# Missing data handling 

## missing count
```{r}
# none of proteins has valid value in all replicates 
# df_missing_normalized %>% drop_na() %>% summarise(num_of_proteins = n())

df_missing_normalized %>% summarise(num_of_missing = n())
```



## simple imputation: replace NA as 0
```{r}
# df_missing_normalized <- df_missing_normalized %>% mutate_all(~replace(., is.na(.), 0))
df_missing_normalized_matrix[is.na(df_missing_normalized_matrix)] <- 0
```


## remove replicates that library size is 0
```{r}
zero_library_filter <- apply(df_missing_normalized_matrix, 2, sum) != 0
df_missing_normalized_matrix <- df_missing_normalized_matrix[, zero_library_filter]
annotation <- annotation[zero_library_filter, ]
```

## filter
```{r}
df_missing_normalized_matrix <- df_missing_normalized_matrix[apply(df_missing_normalized_matrix, 1, function(c) sum(c!=0) >= 5), ]
```



# create DGEList data class
```{r}
group <- factor(annotation$simplified_cancer_type)
y <- DGEList(df_missing_normalized_matrix, group = group)
```


<!-- # Filtering -->
<!-- Should I filter protein with too less expression? -->
<!-- prefer not to filter now cause missing proteins are too sparse -->



# normalization
If use median normalization, correction factor should be feed back to the model
```{r}

```



# classic edgeR estimate overdispersion parameters
```{r}
design <- model.matrix(~group)
colnames(design) <- str_replace(colnames(design), 'group', '')
y <- estimateDisp(y, design = design)

fit <- glmQLFit(y, design)
```


# ANOVA test
```{r}
ANOVA_test <- glmQLFTest(fit, coef=as.character(unique(group))[-1])
```

```{r}
test <- glmQLFTest(fit, coef=c('Colon'))
```


```{r}
ANOVA_DEPs <- topTags(ANOVA_test, 1000, p.value = 0.05)
DEPs <- topTags(test, 1000, p.value = 0.05)
```

```{r}
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# 
# BiocManager::install("GO.db")

goana(test)
```
```{r}
plotMDS(y, labels = group, col=group)
```

```{r}
plotBCV(y)
```


```{r}
plotMD(test)
abline(h=c(-1, 1), col="blue")
```



# Visulation

## Heatmap for DE proteins in ANOVA test
```{r}
# heatmap_dat <- cpm(y, normalized.lib.sizes = TRUE, log=TRUE, prior.count = 2)
```

```{r}
df_missing_normalized_matrix <- log2(df_missing_normalized_matrix + 1)


# annotation_col is a dataframe with rownames corresponding to the colnames in the feeded matrix
pheatmap(df_missing_normalized_matrix[rownames(ANOVA_DEPs$table), ],
         #fontsize_col = 5, 
         #fontsize_row = 4,
         show_rownames = F,
         show_colnames = F,
         main = "Surface", 
         cluster_cols = F,
         cluster_rows = T,
         annotation= data.frame(annotation[1], row.names = annotation$full_name))
```

```{r}
  # if (!requireNamespace('BiocManager', quietly = TRUE))
  #   install.packages('BiocManager')
  # 
  # BiocManager::install('EnhancedVolcano')


 EnhancedVolcano(test$table,
    lab = rownames(test$table),
    x = 'logFC',
    y = 'PValue')
```

# Venn diagram
```{r}

```












# other processing
1. collapse cancer type category
2. fold-change threshold
3. need corresponding gene name for GO and KEGG analysis






# concerns:
imputation missing value with 0 may not be correct