---
title: "raw_data_analysis"
output: html_document
date: "2022-09-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
```



# section 1 data processing
```{r}
nasf_tce_tbl <- read.csv('data/cell line raw data/nasf_tce.csv', row.names = 1)
```


IP0986_1701       Gastric
IP0987_1701       Gastric
IP0988_1701       Gastric
IP0989_1701       Gastric
IP0991_1701       Gastric
IP0992_1701       Gastric
IP7101_1701       Gastric
IP7129_1701       Gastric
IP7128_1701       Gastric
```{r}
gastric_cell_line <- c('IP0986_1701', 'IP0987_1701', 'IP0988_1701',
  'IP0989_1701', 'IP0991_1701', 'IP0992_1701', 'IP7101_1701', 'IP7129_1701', 'IP7128_1701')
tce_nasf <- nasf_tce_tbl %>% select(all_of(gastric_cell_line))
tce_nasf_mat <- as.matrix(tce_nasf)
```

```{r}
omics_metadata <- read_csv('data/support data/omics_metadata.csv')
```


```{r}
# accession_gene_symbol <- data.frame(gene_symbol = accession_metadata$gene,row.names = accession_metadata$accession)
# rownames(tce_nasf) <- accession_gene_symbol[rownames(tce_nasf),]
tce_nasf_gene_level <- rownames_mapping(tce_nasf_mat, omics_metadata, map_from = 'accession_id', map_to = 'gene_symbol')
# use sum or average? ?? sum seems to make more sense
temp <- as_tibble(tce_nasf_gene_level, rownames='symbol') %>% group_by(symbol) %>% summarise_all(sum)
tce_nasf_gene_level <- as.matrix(temp[-1])
rownames(tce_nasf_gene_level) <- temp$symbol
head(tce_nasf_gene_level)
nrow(tce_nasf_gene_level)
```


```{r}
load('cache/count_TPM.rda')
count_TPM_gene_symbol <- rownames_mapping(count_TPM_mat, omics_metadata, map_from = 'ensembl_gene_id', map_to = 'gene_symbol')
# change the column name the same as in spectral count matrix
colnames(count_TPM_gene_symbol) <- c('IP7103_1701', 'IP0972_1701', 'IP0995_1701', 'IP0982_1701', 'IP0999_1701', 'IP7100_1701', 'IP0993_1701', 'IP0981_1701', 'IP7105_1701')
# change the order of column the same as in spectral count matrix
count_TPM_gene_symbol <- count_TPM_gene_symbol[, c('IP0993_1701', 'IP0995_1701', 'IP0999_1701', 'IP0972_1701', 'IP0981_1701', 'IP0982_1701', 'IP7103_1701', 'IP7105_1701', 'IP7100_1701')]
head(count_TPM_gene_symbol)
nrow(count_TPM_gene_symbol)
```



```{r}
# percentage of gene that can be found to have protein product
mean(rownames(count_TPM_gene_symbol) %in% rownames(tce_nasf_gene_level))
# percentage of proteins that have transcription level evidence
mean(rownames(tce_nasf_gene_level) %in% rownames(count_TPM_gene_symbol))
```


! NA %in% NA is treated as 0 in R so need to remove it before use to get the intersection
of gene symbols between two data
```{r}
both <- na.omit(intersect(rownames(count_TPM_gene_symbol), rownames(tce_nasf_gene_level)))
length(both)
```


```{r}
count_TPM_both <- count_TPM_gene_symbol[rownames(count_TPM_gene_symbol) %in% both,]
count_nasf_both <- tce_nasf_gene_level[rownames(tce_nasf_gene_level) %in% both,]
```


reorder row to make the two matrix have the same row order
```{r}
count_nasf_both <- count_nasf_both[both,]
count_TPM_both <- count_TPM_both[both, ]
head(count_nasf_both)
head(count_TPM_both)
```

```{r}
pheatmap::pheatmap(log(count_TPM_both+1),
                   cluster_rows = TRUE,
                   cluster_cols = FALSE,
                   show_rownames = FALSE
                   )
```



```{r}
pheatmap::pheatmap(log(count_nasf_both+1),
                   cluster_rows = TRUE,
                   cluster_cols = FALSE,
                   show_rownames = FALSE
                   )
```



# missing proteins
check how many missing proteins is expressed in the 8 gastric sample in the protein data and in RNA-seq data  
```{r}
missing_protein_df <- readxl::read_xlsx('data/support data/PE2-5.xlsx')
missing_protein_df <- missing_protein_df[-c(1:12), ]

mp_in_gene_symbol <- missing_protein_df$`gene name(s)`
sum(rownames(tce_nasf_gene_level) %in% mp_in_gene_symbol)
sum(both %in% mp_in_gene_symbol)
```


There two many gene, so let's only keep missing proteins
```{r}
count_nasf_missing <- count_nasf_both[rownames(count_nasf_both) %in% mp_in_gene_symbol, ]
count_TPM_missing <- count_TPM_both[rownames(count_TPM_both) %in% mp_in_gene_symbol, ]
```

```{r}
pheatmap::pheatmap(log(count_TPM_missing+1),
                   cluster_rows = FALSE,
                   cluster_cols = FALSE,
                   show_rownames = FALSE
                   )
```

```{r}
pheatmap::pheatmap(log(count_nasf_missing+1),
                   cluster_rows = FALSE,
                   cluster_cols = FALSE,
                   show_rownames = FALSE
                   )
```

# check paired correlation


# scatterplot
```{r}
matplot(log(count_nasf_missing+1), log(count_TPM_missing+1))
```


```{r}
y <- log(rowSums(count_nasf_missing)+1)
x <- log(rowSums(count_TPM_missing)+1)
plot(log(rowSums(count_nasf_missing)+1), log(rowSums(count_TPM_missing)+1))
lm()
```


```{r}
library(corrplot)
corrplot(cor(count_nasf_missing, count_TPM_missing), method='color')
```

```{r}
corr <- cor(t(count_nasf_missing), t(count_TPM_missing))
mean(diag(corr) > 0.2, na.rm = T)
library(corrplot)
corrplot(corr, method = 'color', na.label.col='black')
```

