---
title: "analysis"
output: html_document
date: "2022-07-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# 
# BiocManager::install("DEqMS")
```

```{r}
library(DEqMS)
library(tidyverse)
library(pheatmap)
```


# Reading data
```{r}
SpC <- read.delim("data/CellLineCompartment_TCE_SumTickNormalization.tsv", sep = '\t', skip=2, row.names = 'Accession')
# annotation <- read_tsv("data/CellLineCompartment_TCE_SumTickNormalization.tsv", n_max=3, col_select = c(-1), col_names = F, show_col_types = FALSE)
```

## create column annotation
```{r}
disease_fulltype <- str_extract(colnames(SpC), '.+(?=\\.IP)')
disease <- disease_fulltype
disease[str_detect(disease_fulltype, 'Breast')] <- 'Breast'
disease[str_detect(disease_fulltype, 'Leukemia')] <- 'Leukemia'
IPAS_annotation <- data.frame(disease = disease, row.names = colnames(SpC))
# should I use Pediatric GVHD as control?
unique(disease)
```


## read-in newest missing protein list
```{r}
missing_protein_df <- readxl::read_xlsx('data/PE2-5.xlsx')
missing_protein_df <- missing_protein_df[-c(1:12), ]
head(missing_protein_df)
```



#===============================================================================
# Analysis

```{r}
y <- SpC_List(SpC, IPAS_annotation, NA_substitution = 0, proteins_filter = missing_protein_df$Accession)
```


```{r}
basic_describe(y)
```


# collapse 'Lymphoblast', 'Neuroblastoma', 'Prostate' since the group size is too small for reliable test 
```{r}
IPAS_annotation <- IPAS_annotation %>% mutate(disease = if_else(disease %in% c('Melanoma', 'Lymphoblast', 'Neuroblastoma', 'Prostate'), 'Other', disease))
```


# filter the latest missing proteins and outlier replicate
The sample Prostate IP0872_1701 H660  has SpC counts of 55539 almost 2 magnitude larger than any other sample
treat it as an outlier
after filtering, the number of missing proteins reduce from 1856 -> 1581
```{r}
y <- SpC_List(SpC, IPAS_annotation, NA_substitution = 0, proteins_filter = missing_protein_df$Accession, replicates_remove = 'Prostate.IP0872_1701.H660')
```


```{r}
basic_describe(y)
```

<!-- # Data pre-filter -->
<!-- Since the data is normalized not raw data, not use a pre-filter here -->
<!-- <!-- remove proteins that is not confidentially detected --> -->
<!-- ```{r} -->
<!-- # N <- min(table(IPAS_annotation$disease)) -->
<!-- # keep <- rowSums(SpC_matrix > 1) >= N -->
<!-- # SpC_matrix_dense <- SpC_matrix[keep,] -->
<!-- # SpC_matrix_dense <- SpC_matrix -->
<!-- ``` -->


perform test
```{r}
res_lst <- one_vs_all_test(y, logFC = 1, PAjusted = 0.05)
```


res_lst
```{r}
unique_DEPs_TCE <- unique(unlist(lapply(res_lst, rownames)))
# paste0(unique_DEPs, collapse = '\n')
```



# Heatmap for the entire data matrix
```{r}
plot_heat_map(y, upper_limit = 5, row_keep = unique_DEPs, show_rownames = F, main='Cell Component Media')
```





# try use venn diagram to show uniquely missing DEPs in each dataset
```{r}
library(VennDiagram)
venn.diagram(
  x = list('TCE'=unique_DEPs_TCE, 'exosome'=unique_DEPs_exosome, 'media'=unique_DEPs_media, 'nucleas'=unique_DEPs_nucleas),
  filename = 'output/venn_diagramm.png',
  output=T
)
```


 Breast         Colon       Gastric        Glioma      Leukemia LungSmallCell   Lymphoblast      Melanoma Neuroblastoma         NSCLC       Ovarian 
   39            10             9            10            15            27             2             7             2            60            15 
   Pancreatic      Prostate 
           29             5 
```{r}
# if (!require(devtools)) install.packages("devtools")
# devtools::install_github("yanlinlin82/ggvenn")
library(ggvenn)

unique_proteins_subsets <- lapply(res_lst, rownames)
# check number of missing proteins in each cancer type
# lapply(unique_proteins_subsets, length)

ggvenn(
  unique_proteins_subsets[c('NSCLC', 'Breast', 'Pancreatic', 'LungSmallCell')], 
  fill_color = c("#0073C2FF", "#EFC000FF", "#868686FF", "#CD534CFF"),
  stroke_size = 0.5, set_name_size = 7
  )
```

```{r}
library(RColorBrewer)

breaksList = seq(0, 5, by = 0.05)

# annotation_col is a dataframe with rownames corresponding to the colnames in the feeded matrix
pheatmap(log2(y$matrix[unique_DEPs, ] + 1),
         color = colorRampPalette(colors = c("black", "yellow", "darkorange"))(length(breaksList)),
         breaks = breaksList,
         #fontsize_col = 5, 
         #fontsize_row = 4,
         show_rownames = F,
         show_colnames = F,
         cluster_cols =  F,
         cluster_rows = T,
         annotation_col=IPAS_annotation)
```



PCA
```{r}
library(ggfortify)
pca_res <- prcomp(t(SpC_matrix_dense[unique_DEPs, ]))
autoplot(pca_res, data = IPAS_annotation, colour = 'disease')
SpC_matrix_dense[unique_DEPs, ]
```



```{r}
missing_DEPs_info <- missing_protein_df %>% filter(Accession %in% unique_DEPs)
```



```{r}
pie(table(missing_DEPs_info$PE))
```





